TRPγ regulates lipid metabolism through Dh44 neuroendocrine cells

Version 1

Abstract

Abstract

Understanding how the brain controls nutrient storage is pivotal. Transient receptor potential (TRP) channels are conserved from insects to humans. They serve in detecting environmental shifts and in acting as internal sensors. Here, we found that a TRPγ mutant exhibited in Drosophila melanogaster are required for maintaining normal lipid and protein levels. In animals, lipogenesis and lipolysis control lipid levels in response to food availability. Lipids are mostly stored as triacylglycerol in the fat bodies (FBs) of D. melanogaster. Interestingly, trpγ deficient mutants exhibited elevated TAG levels and our genetic data indicated that Dh44 neurons are indispensable for normal lipid storage but not protein storage. The trpγ mutants also exhibited reduced starvation resistance, which was attributed to insufficient lipolysis in the FBs. This could be mitigated by administering lipase or metformin orally, indicating a potential treatment pathway. Gene expression analysis indicated that trpγ knockout downregulated brummer, a key lipolytic gene, resulting in chronic lipolytic deficits in the gut and other fat tissues. The study also highlighted the role of specific proteins, including neuropeptide DH44 and its receptor DH44R2 in lipid regulation. Our findings provide insight into the broader question of how the brain regulates nutrient storage.

Introduction

Nutrient storage is not a simple matter of accumulation but involves a sophisticated network of signals and receptors, feedback loops, and hormonal influences, all choreographed by the brain. The mechanisms involve various tissues and organs, including the liver, muscles, and adipose tissue, each playing a distinct role in the storage and release of vital nutrients like glucose, fats, and proteins in mammals 1. Key to this regulation are specialized cells and molecular pathways that respond to dietary intake, energy expenditure, and changes in the internal and external environment 2. For instance, the brain receives and interprets signals about the body’s energy status through hormones such as insulin and leptin, and nutrients themselves, which inform decisions on whether to store or mobilize energy reserves in mammals 3,4. Moreover, the role of Transient Receptor Potential (TRP) channels as molecular sensors that detect changes in the environment and internal metabolic status highlights the level of molecular sophistication involved in nutrient storage and energy balance 5–8. These channels bridge external stimuli with internal responses, ensuring the body adapts efficiently to maintain homeostasis 9. This concert of mechanisms, from the molecular to the systemic level, exemplifies the elegance with which the brain ensures our energy needs are met in a changing environment, highlighting the intricate interplay of biological process that sustains life 10. Understanding these processes opens up new vistas in our approach to diet, disease management, and overall health optimization.

The use of a genetically tractable model organism such as Drosophila melanogaster could provide fundamental insights into the potential relationship between TRP channels and metabolic syndrome. This study delves into the function of TRPγ channels, specifically their role in managing lipid and protein levels. Through genetic manipulation, we explored the consequences of TRPγ mutations on the metabolic processes, revealing critical insights into the regulation of lipid storage and breakdown. Our investigation sheds light on the balance between lipogenesis and lipolysis, processes that respectively synthesize and degrade lipids based on the availability of food, and how these processes are affected by TRPγ activity. By examining the metabolic anomalies in trpγ deficient mutants, including alterations in lipid storage and the consequent impact on starvation resistance, this study advances our understanding of metabolic regulation. Furthermore, we uncover the potential therapeutic benefits of lipase and metformin in counteracting the metabolic deficiencies caused by TRPγ mutations. Through detailed gene expression analysis, we identify the downregulation of crucial genes involved in lipid metabolism, offering new perspectives on the genetic underpinnings of nutrient storage regulation. The implications of this research are far-reaching, not only enhancing our comprehension of basic biological functions but also providing avenues for therapeutic interventions in metabolic disorders.

Results

TRPγ mutants exhibit reduced sugar level alongside elevated lipid and protein levels

In order to understand how D. melanogaster regulates the maintenance of major nutrient levels, including carbohydrates, lipids, and proteins, TRP channel mutants were examined as potential candidates. A recent study revealed that among these mutants, the trpγ mutant displayed only reduced carbohydrate levels 11, suggesting the involvement of additional TRP channels in regulating major nutrients. D. melanogaster possesses 13 members of the TRP channel family. Flies with mutations in trpM, trpML, and nompC showed high mortality rates with minimal survival, whereas those with mutations in the remaining 10 TRP superfamily genes were viable and healthy when homozygous. Investigation into the lipid and protein levels of other available TRP superfamily mutants revealed that, apart from the trpγ mutant, all tested mutants exhibited normal levels (Figure 1A and B). Specifically, the trpγ1 mutant showed 1.3 times higher triacylglycerol (TAG) levels and 1.5 times higher protein levels compared to the control. These metabolic changes were unique and specific to the trpγ mutants among the evaluated TRP superfamily mutants. Previous observations indicated that trpγ mutants had lower cellular sugar levels and stored glycogen. Consistently, trpγ mutants exhibited reduced levels of cellular sugars (glucose and trehalose), stored glycogen, and sugar levels in the hemolymph (Figure 1C–E). In conclusion, it was found that trpγ1 mutants displayed lower carbohydrate levels but higher lipid and protein levels compared to the control. Importantly, the increased TAG and protein levels observed in trpγ1 mutants were confirmed with a second trpγ1 allele (trpγG4) (Figure 1F and G). Furthermore, we restored the deficiency in lipid and protein levels using a genomic trpγ+ transgene, g(trpγ), and by introducing a UAS-trpγ cDNA, controlled by GAL4, into the trpγ locus (trpγG4 flies; +).

TRPg mutants exhibit altered carbohydrate, lipid, and protein levels. (A) TAG level measurement in the whole-body extract from control (w1118) and TRP channel mutant lines (n=9). (B) Protein level measurement in the whole-body extract from control (w1118) and TRP channel mutant lines (n=6). (C) Measurement of total glucose and trehalose levels (µg/mg) in the whole-body extracts of control (w1118) and trpg1 adult males (n=6). (D) Measurement of tissue glycogen levels (µg/mg) in adult control (w1118) and trpg1 males (n=6). (E) Measurement of Hemolymph glucose + trehalose level in the male flies of control (w1118) and trpγ1 (n=6). (F) Measurement of TAG level in adult males to test for rescue of the TAG defect in trpg flies with the UAS-trpg and the trpgG4 or with the trpg genomic transgene (n=6-10). (G) Rescue experiments showing the measurement of protein levels in adult flies with the indicated genotypes (n=6).

Comparisons between multiple experimental groups were conducted via single-factor ANOVA coupled with Scheffe’s post hoc test. The asterisks indicate significant differences from the controls (**P<0.01).

TRPγ regulates lipid metabolism through Dh44 neuroendocrine cells in the pars intercerebralis

We recently proposed that TRPγ expression in the six DH44 neuroendocrine cells in the pars intercerebralis (PI) located in the dorsal medial area of the brain is essential to maintain normal carbohydrate levels in tissues (Figure 2–figure supplement 1A1–A3) 11. Therefore, we next sought to determine whether protein and lipid levels can be regulated by the same neuroendocrine cells. To test this, we used a transgenic fly encoding an inwardly rectifying K+ channel (UAS-Kir2.1) to inactivate specific neurons (Figure 2A and B) 12. Interestingly, inactivating the Dh44 neurons significantly increased the lipid levels but not protein levels (Figure 2A and B). Approximately 16 neurons in the PI express dILP2, which are also overlapped by two DH44-positive neurons in 5 to 10 days old adult male flies (Figure 2–figure supplement 1B1–B3) 13. However, inactivation of dILP2 neurons did not impair either lipid or protein levels (Figure 2A and B). In conclusion, our findings indicated that DH44 neuroendocrine cells contribute to lipid regulation but not protein regulation.

Dh44 neurons are essential for regulating lipid tissue levels. (A) Tissue TAG level measurement in whole-body extracts of adult male flies after silencing of Dh44-GAL4 and dILP2-GAL4 with UAS-Kir2.1 (n=6). (B) Measurement of tissue protein level in whole-body extracts of adult male flies. Brain-specific Dh44-GAL4 and dILP2-GAL4 neurons were ablated using UAS-Kir2.1 (n=6). (C) Measurement of TAG from whole-body extract of adult male flies in the indicated genotypes (n=6). (D) Measurement of tissue protein level from the whole-body extracts of adult males (n=6). (E) TAG levels in whole-body extracts after RNAi knockdown of trpγ mutants under control of the Dh44-GAL4 (n=6). (F) Measurement of area (µm2) of LDs in adult fat body across the indicated genotypes involved the selection of the 30 largest LDs, choosing the top 10 LDs from each sample for analysis (n=3). (G–L) Nile red stating of fat body from the male of indicated genotypes. The scale bar represents 50 µm.

All values are reported as means ± SEM. Comparisons between multiple experimental groups were conducted via single-factor ANOVA coupled with Scheffe’s post hoc test. The asterisks indicate significant differences from the controls (*P<0.05, **P<0.01). Each dot indicates the distribution of individual sample values. (+) and (-) indicate the presence or absence of the indicated transgenes, respectively.

To assess whether trpγ expression in Dh44 neurons is sufficient to restore normal lipid and protein levels in trpγ mutant flies, we expressed the UAS-trpγ under the control of Dh44-GAL4 in the trpγ1 mutant background (Figure 2C and D). Our findings indicated that the increased lipid levels in the trpγ1 mutant background decreased to normal levels via the expression of the trpγ transgene in the DH44 neuroendocrine cells but not in its parent strains (Figure 2C). Again, the expression of the trpγ transgene in the dILP2 neurons had no appreciable effects (Figure 2C). In contrast, the expression of the trpγ transgene did not decrease the protein levels in either the DH44 or the dILP2 neurons (Figure 2D). This indicates that TRPγ is sufficient for the regulation of lipid levels in the DH44 cells. Next, RNAi-mediated knockdown experiments were conducted to further examine the role of trpγ in Dh44 neurons. Interestingly, trpγ knockdown in Dh44 neurons significantly increased lipid levels, whereas flies harboring only the Dh44-GAL4 or the UAS-trpγRNAi transgenes displayed normal lipid levels (Figure 2E). This indicated that TRPγ is needed for the regulation of lipid levels. Furthermore, the total TAG level in the trpγ1 flies was higher in both males and females (Figure 2–figure supplement 1C), meaning that the functions of trpγ are not sex-specific.

In D. melanogaster, lipids are mainly stored in the form of TAG and cholesterol ester in the adipose tissue [i.e., fat bodies (FBs)] as lipid droplets (LDs) 14. Consistent with the increased TAG in tissues, Nile red staining of the FBs of trpγ1 and trpγG4 flies exhibited larger lipid mass compared to control animals (Figure 2F–I). These LDs returned to their normal size through the expression of trpγ in Dh44 neurons but not dILP2 neurons (Figure 2F and 2J–L). Additionally, inactivating the Dh44 cells (Dh44-GAL4/UAS-Kir2.1) recapitulated the enlarged LD phenotype of the trpγ mutant flies (Figure 2–figure supplement 1D–G). Overall, our findings indicated that trpγ expression regulates lipid and carbohydrate homeostasis but not protein levels in DH44 neuroendocrine cells in the PI.

TRPγ mutants exhibit starvation susceptibility and deficits in lipolysis

Higher lipid cellular levels may decrease lifespan 15,16. Therefore, we measured the lifespan of control and trpγ1 flies fed with a standard cornmeal diet (Figure 3A). However, the lifespans of control and trpγ1 flies were not significantly different under normal conditions. The LT50 of the control was 59.34 ± 0.92 days and that of trpγ1 was 55.75 ± 2.35 days. In contrast, trpγ appeared to be required for proper metabolism under starvation conditions, as demonstrated by the decreased starvation resistance of the trpγ1 and trpγRNAi knockdown flies in Dh44 cells (Figure 3B) 11. This defect was fully recovered by the expression of UAS-trpγ under the control of Dh44-GAL4 (Figure 3B).

trpγ1 have deficits in the lipolytic pathway under starvation conditions. (A) Survival assay to measure the total survival time (days) of control (w1118) and trpγ1 male flies fed with a normal corn meal diet (n=8). (B) Survival assay to measure the survival time (h) of the indicated genotypes with male flies under starvation conditions (n=4). (C) TAG level measurement in control (w1118) and trpγ1 adult male flies in both sated (0 h starvation) and starved (24 h starvation) conditions (n=8). (D–E) Nile red staining of the LDs extracted from FB of (D) w1118 and (E) trpγ1 flies under sated (D1 and E1) and starved (D2 and E2) conditions, respectively. The scale bar represents 50 µm. (F) Measurement of area (µm2) of the LDs extracted from the FBs of w1118and trpγ1 flies under sated (0 h starvation) and starved (24 h starvation) conditions (n=3). (G) Measurement of TAG level with controls and the flies after expressing UAS-AMPKTD under the control of Dh44-GAL4 in the trpγ1 mutant background (n=4).

All values are reported as means ± SEM. Survival curves in A and B were estimated for each group, using a Kaplan-Meier method and compared statistically using the log-rank tests. Comparisons between multiple experimental groups in C, F, G, and H were conducted via single-factor ANOVA coupled with Scheffe’s post hoc test. The asterisks indicate significant differences from the controls (**P<0.01).

The starvation sensitive phenotype in the trpγ mutants may have been due to decreased carbohydrate storage in tissues, including glucose, trehalose, and glycogen. However, we previously demonstrated that trpγ mutants could utilize carbohydrates under starved conditions 11. Furthermore, the elevated whole-body TAG levels in the trpγ mutant flies might suggest that they were unable to break down stored lipids even under starving conditions. Metazoans must coordinate the metabolism of glycogen, lipid, and protein to maintain metabolic homeostasis during fasting periods, thus providing an appropriate energy supply across tissues. Therefore, the total TAG levels of the control and trpγ mutant flies were assessed under sated and starved (starvation for 24 h) conditions (Figure 3C–F). When the control flies were deprived of food for 24 h, their TAG levels decreased significantly (Figure 3C). In contrast, the trpγ1 mutants exhibited no changes in whole-body TAG levels before and after starvation (Figure 3C). To further confirm that lipolysis was restricted in the trpγ mutants under starvation conditions, LDs were stained, after which we measured the sizes of individual LDs accumulated throughout the whole fat bodies of the flies (Figure 3D–F). Despite considerable variations, the sizes of the LDs in the control flies were significantly reduced after starvation. In contrast, no significant differences in LD sizes were identified between the sated and starved conditions in the trpγ1 flies. We examined how overexpression and knockdown of trpγ in Dh44 neurons affect the starvation phenotype. Overexpressing trpγ in Dh44 cells resulted in similarity to the wild-type in both sated and starved conditions, as well as normal survival time under starvation conditions (Figure 3B and Figure 3–figure supplement 1A). Conversely, trpγRNAi knockdown flies in Dh44 neurons reproduced phenotypic traits observed in TRPγ mutants, including decreased lipid levels (Figure 3–figure supplement 1B) and reduced survival time under starvation conditions (Figure 3B).

Adenosine monophosphate-activated protein kinase (AMPK) serves as the master controller for maintaining energy balance in cells, coordinating metabolic pathways 17. It regulates the balance between building up and breaking down substances to ensure cellular stability during metabolic stress. AMPK is a key target for treating metabolic diseases like type 2 diabetes and obesity, as its activation increases fatty acid oxidation 18. We wonder if activating AMPK in Dh44 neurons serves as a rescue mechanism for lipolysis. The expression of AMPKTD, activated form of AMPK 19 in Dh44 neurons indeed restored the elevated TAG levels observed in trpγ1 (Figure 3G). This suggests that AMPK functions as a downstream component of Dh44 neuronal activation.

Recovery of the starvation resistance of the TRPγ mutant via metformin treatment

Metformin is widely used to treat many metabolic diseases such as type II diabetes 20. We previously demonstrated that 1 to 5 mM metformin can induce hypoglycemia, in addition to suppressing fat storage in flies 21. Therefore, we next sought to test whether the increased lipid levels of the trpγ mutant can be reduced to a normal level via oral administration of metformin. First, we measured TAG levels at 0, 7, and 14 days after dietary administration of 1 mM or 5 mM of metformin (the treatments were prepared by mixing the appropriate metformin concentrations into standard cornmeal diets) (Figure 4A and B). The high TAG levels in the trpγ mutant (41.91 ±4.56) were decreased to levels similar to those of the control at 7 and 14 days (30.03 ±4.34 and 33.79 ±3.20, respectively) after treatment with 1 mM metformin (Figure 4A). Furthermore, although oral administration of 1 mM metformin did not affect the TAG level in the control, the 5 mM metformin treatment significantly reduced the TAG level of the control flies at 14 days (Figure 4B). Moreover, the 5 mM metformin treatment was more effective in reducing the TAG level in the trpγ mutant. These findings indicated that oral administration of metformin can either suppress lipogenesis or enhance lipolysis in D. melanogaster. To further confirm the effects of metformin treatment, LDs were analyzed under the same condition (Figure 4C and D). The sizes of the LDs in the control and trpγ mutant flies were significantly reduced after treatment with 1 mM metformin at 14 days or 5 mM metformin at 7 and 14 days (Figure 4D).

Rescue of starvation susceptibility phenotype using the lipolytic drug metformin. (A) TAG level measurement at 0, 7, and 14 days in the control (w1118) and trpγ1 adult male flies after dietary exposure to 1 mM metformin (n=6). (B) TAG measurement at 0, 7, and 14 days in w1118 and trpγ1 adult male flies after dietary exposure to 5 mM metformin (n=6). (C) Pictures of Nile red staining of LDs after dietary exposure to 1 mM and 5 mM metformin in standard fly food for 7 and 14 days in w1118 and trpγ1 flies. The scale bars represent 50 µm. (D) Measurement of area (µm2) of LDs extracted from the FB of w1118and trpγ1 flies after dietary exposure to 1mM and 5 mM metformin in standard fly food for 7 and 14 days (n=3). (E) Survival assay to measure the survival time (h) of control (w1118) and trpγ1 males after dietary exposure to 1 mM metformin in 1% agar food (n=6). (F) Starvation survival assay of control (w1118) and trpγ1 males after dietary exposure to 5 mM metformin in 1% agar food (n=6).

All values are reported as means ± SEM. Comparisons between multiple experimental groups were conducted via single-factor ANOVA coupled with Scheffe’s post hoc test. Each dot indicates the distribution of individual sample values. Survival curves in E and F were estimated for each group, using a Kaplan-Meier method and compared statistically using the log-rank tests. The asterisks indicate significant differences from the controls (**P<0.01).

Next, we sought to assess whether metformin could increase starvation resistance in the control and trpγ mutant flies. To test this hypothesis, the starvation resistance of the flies was measured with or without metformin. Neither of the metformin concentrations appeared to affect the starvation resistance of the control flies (Figure 4E and F). However, although metformin did not restore the starvation resistance of the trpγ mutant flies to the level of the controls, it did significantly extend their survival (Figure 4E and F; LT50 of the control at 1 mM and 5 mM: 56.06 ±1.40 h and 52.39 ±2.09 h, respectively; LT50 of the trpγ1 mutant at 1 mM and 5 mM: 50.35 ±3.34 h and 46.42 ±1.54 h, respectively). Our findings thus demonstrated that oral metformin administration can rescue TRPγ-mediated metabolic syndrome.

Recovery of the TRPγ mutant through oral lipase administration or lipid absorption

The intestine is a key organ for lipid absorption and metabolism 22 and disturbances in intestinal lipid metabolism are associated with hyperlipidemia 23. Therefore, we next sought to determine whether trpγ mutation reduces lipase production in the abdomen by comparing lipid accumulation in the intestine. The Drosophila midgut is divided into R1–R5 regions, each with distinct functions in nutrient processing and absorption. Notably, the R2 region serves as a critical interface between the fly’s diet and metabolic processes 24. To assess lipid distribution, we conducted Nile red staining in control and mutant animals (Figure 5A–D). Mutant flies displayed elevated lipid deposits on the intestinal wall of the R2 region (Figure 5B), a phenotype restored to normal levels by expressing trpγ in its own cells or Dh44 neurons (Figure 5C–E). Consequently, we quantified lipid droplets specifically within the intestinal wall of the R2 region in trpγ1 flies, with no such accumulation observed in control or rescued flies (Figure 5E). These findings indicate that TRPγ plays a crucial role in the brain-gut axis, specifically in controlling lipid metabolism in the intestine, as evidenced by the lipid accumulation observed in the gut region of trpγ1. To clarify if our observations were due to insufficient activity of lipase in the gut, lipase enzyme mixed with 1% agar food was supplied to both control and mutants flies. Given that the brief activity of lipase in the intestine, lipase-containing agar food was provided to the flies every 12 h. Surprisingly, supplementation with lipase, but not denatured lipase, under starvation conditions increased the survival of trpγ mutant flies to levels comparable to those of the control flies (Figure 5F; LT50 of lipase treated-control and trpγ1: 50.35 ±0.93 h and 44.94 ±1.33 h, respectively). To support the hypothesis that the extended lifespan of mutant flies during starvation is associated with dietary lipase intake, we examined lipid accumulation (Figure 5–figure supplement 1A–E). Lipase treatment resulted in the breakdown of lipid deposits in the mutant R2 region of intestine. Importantly, only active lipase showed an enhanced survival time under starvation conditions in mutant flies, while denatured lipase did not exhibit this effect. Therefore, our findings suggest that lipase may play a role in digesting accumulated lipid in the intestine, though the mechanisms of lipase absorption remain unclear.

Dietary supply of lipase, glycerol, mixed TAG, and free FA to rescue the starvation sensitivity phenotype. (A–D) Nile red staining of LDs in full gut and the magnified R2 region of adult male flies. (A) w1118, (B) trpγ1, (C) trpγ1,UAS-trpγ/trpγG4, (D) trpγ1,UAS-trpγ/trpγ1;Dh44-GAL4/+. The scale bar represents 50 µm. The arrow indicates the orientation of intestine from anterior to posterior. (E) Measurement of area of LDs in the R2 region of trpγ1 (n=10). Note that control and rescued flies have no LD. (F) Survival assays of w1118 and trpγ1 flies under starvation condition by feeding 0.1% lipase or 0.1% denatured lipase mixed into 1% agar food (n=8). (G) Starvation survival assay to measure the survival time (h) of w1118 and trpγ1 male flies after feeding 1% glycerol mixed into 1% agar food (n=4). (H) Survival assay to measure the survival time (h) of w1118 and trpγ1 flies after feeding 0.2% hexanoic acid (HA) supplemented into 1% agar food (n=8). (I) Survival assay to measure the survival time (h) of w1118 and trpγ1 flies after feeding 0.2% mixed (mono-, di-, and tri-) glycerides (n=4).

All values are reported as means ± SEM. Survival curves were estimated for each group, using a Kaplan-Meier method and compared statistically using the log-rank tests. (**P<0.01).

Next, we sought to assess the involvement of each of the components of TAG in the recovery of starvation susceptibility. TAG is composed of glycerol and three fatty acids. Therefore, we first fed the flies with 1% glycerol only. Glycerol extended the survival of the control and trpγ1 flies by 3–4-fold (Figure 5G). However, the LT50 of the control and trpγ1 were still significantly different (Figure 5H; LT50 of control and trpγ1: 244.5 ± 20.10 h and 117.75 ± 3.32 h, respectively). Next, the flies were fed with 0.2% and 0.5% hexanoic acid (HA) (Figure 5H and Figure 5–figure supplement 1F; 0.2% HA, LT50 of control and trpγ1: 65.44 ± 1.94 h and 52.32 ±2.05 h, respectively; 0.5% HA, LT50 of control and trpγ1: 56.27 ±1.30 h and 46.27 ±1.47 h, respectively). Additionally, we also tested the effects of 0.2% and 0.5% concentrations of a TAG mixture (combination of mono-, di-, and tri- acylglycerol) (Figure 5I and Figure 5–figure supplement 1G; 0.2% TAG, LT50 of control and trpγ1: 52.93 ±5.12 h and 50.1 ±3.74 h, respectively; 0.5% TAG, LT50 of control and trpγ1: 65.1 ±2.6 h and 57.75 ±3.32 h, respectively). The HA and mixed TAG treatments significantly extended the survival of the trpγ1 mutants, albeit not to levels similar to those of the control. This indicated that the TRPγ mutants could not fully absorb and burn the digested lipid. Furthermore, lipid-only dietary supplementation markedly limited the LT50 of the control. Importantly, supplementation with lipolytic drugs, lipase, TAG, and free fatty acids effectively rescued the survival of trpγ mutants under starvation conditions. This observation was consistent with our finding that trpγ mutants are unable to degrade lipid stores under starvation conditions. The dietary rescue of trpγ mutants suggests that although the trpγ mutation increases the levels of stored lipids, the flies were not able to utilize them when starved. Taken together, these observations indicate that TRPγ plays an important role in maintaining systemic lipid levels through proper expression of triglyceride lipase.

Downregulation of the lipolytic gene brummer in trpγ mutants

The trpγ1 mutants exhibited marked alterations in the levels of major nutrients. Therefore, we next sought to analyze the transcriptional levels of genes related to gluconeogenesis, lipogenesis, and lipolysis. Gluconeogenesis is a metabolic process that produces glucose from non-carbohydrate carbon substrates. The genome of D. melanogaster harbors two gluconeogenic genes: fructose-1,6-bisphosphatase (fbp) and phosphoenolpyruvate carboxykinase 1 (pepck1) 25. The results of our real-time quantitative reverse transcription PCR (qRT-PCR) analyses indicated that the transcription levels of fbp and pepck1 were not significantly different between control flies under sated and starved conditions (Fig. 6A). Next, we analyzed two lipogenic genes: acetyl–CoA carboxylase (acc) and desaturase 1 (desat1) (Figure 6A) 26,27. No significant differences were identified between the control and trpγ1 flies under both conditions. Finally, we investigated the expression of the lipolytic gene brummer (bmm) (Figure 6A) 28. Under starved conditions, the transcriptional level of bmm was highly increased in the control. In contrast, bmm was downregulated in trpγ1 under sated and starved conditions. Therefore, the transcriptional regulation of w1118 and trpγ1 during starvation was markedly and significantly different. Brummer, a homolog of human adipocyte triglyceride lipase (ATGL), is associated with lipid storage in the form of LDs 29,30. These results suggest that the lipolytic process is not sufficient to deplete stored lipids in the trpγ1 mutant, resulting in excessive lipid storage. Therefore, the survival time of the trpγ1 mutant during starvation was significantly shorter than that of the control.

Quantitative analysis of the expression of gluconeogenic, lipogenic, and lipolytic genes and the effect of bmm expression and metformin feeding. (A) qRT-PCR analysis to measure the expression of gluconeogenic genes (fbp, pepck1), lipogenic genes (acc, desat1), and a lipolytic gene (bmm) under sated (0 h starvation) and starved (24 h starvation) conditions in w1118 and trpγ1 flies. The relative fold change in the expression of starvation-induced genes (gluconeogenic genes: fbp and pepck1; lipogenic genes: acc and desat1; and lipolytic gene: bmm) was determined in whole-body samples of male adult w1118 and trpγ1 flies by qRT-PCR. Tubulin was used as an internal control to standardize the samples. Each graph shows the number of evaluated samples (n=3). (B) Fluorescence microscopic imaging of bmm::GFP expression in the FBs of w1118 and trpγ1 under sated (0 h starvation) and starvation (24 h starvation) conditions. The scale bar represent 50 µm. (C) Quantification of intensity level of bmm::GFP in the FBs of w1118and trpγ1 under sated and starvation conditions. (D–H) Nile red staining of lipids in the FB of flies with the indicated genotypes. The scale bar represent 50 µm. (I) Measurement of area of LDs from samples D–H entailed selecting a total of 30 LDs, with the 10 largest LDs chosen from each sample for analysis (n=3). (J) qRT-PCR analysis to measure the expression of acc, desat1, and bmm from the whole body samples of flies after feeding 5 mM metformin for 1 day (n=3).

All values are reported as means ± SEM. Comparisons between multiple experimental groups were conducted via single-factor ANOVA coupled with Scheffe’s post hoc test. The asterisks indicate significant differences from the controls (*P<0.05, **P<0.01).

Transgenic flies expressing bmm::GFP (bmm promoter GFP) in their FB cells were used to examine bmm expression in LDs 30. The fluorescent images obtained from the transgenic flies were then used to confirm our qRT-PCR results. Two images of control flies under sated and starved conditions were first analyzed both qualitatively and quantitatively (Figure 6B and C). Next, trpγ1 was analyzed under the two conditions (Figure 6B and C). A reduction in fluorescent intensity was clearly detectable in the trpγ1 mutant using the transgenic reporter line. These results further confirmed that trpγ1 has aberrant regulation in the lipolytic pathway.

To further substantiate the idea that the reduction in the expression level of bmm is the primary cause of the enlarged LDs in trpγ mutant flies, we measured LDs in the FB after expressing UAS-bmm using specific GAL4 drivers: trpγG4, Dh44-GLA4, r4-GAL4 (fat body specific), Myo1A-GAL4 (gut enterocyte specific), and Dmef-GAL4 (muscle specific). Only trpγG4, Dh44-GAL4 and r4-GAL4 were able to restore the enlarged LDs defects in the FB after expressing UAS-bmm (Figure 6D–I). Overall, these findings suggest that the downregulation of bmm is one of the factors contributing to the elevated lipid defects in trpγ1.

To conclusively explore whether metformin targets enzymes involved in lipid metabolism, specifically lipolytic or lipogenic genes, we administered 5 mM metformin supplemented in normal food for one day and measured the expression levels of bmm, acc, and desat1. Oral supplementation of metformin resulted in an increase in the expression level of the lipolytic gene, bmm lipase of the control as well as trpγ1 (Figure 6J). Importantly, there was no observed effect of metformin on the expression levels of lipogenic genes. This data provides insights into the mode of action of metformin, suggesting a specific impact on lipolysis.

Assessment of the role of DH44 and its receptors (DH44R1 and DH44R2) in relation to the utilization of fat

In Drosophila, two receptors have been identified for DH44 31. In order to investigate the potential roles of DH44 in fat utilization and identify the receptor responsible for nutrient regulation, we examined lipid and protein levels in Dh44Mi, Dh44R1Mi, and Dh44R2Mi. Dh44Mi and Dh44R1Mi showed normal total TAG levels, while Dh44R2Mi exhibited higher lipid levels like trpγ1 (Figure 7A). Additionally, Nile red staining of LDs in the FBs revealed that Dh44Mi and Dh44R2Mimutants possessed larger LDs (Figure 7B–F), although protein levels remained unaffected across all three mutants (Figure 7G). Further investigations quantified the expression of the bmm in the FBs of Dh44Mi, Dh44R1Mi, and Dh44R2Mi under both sated and starved conditions. In sated conditions, bmm expression was found to be downregulated in Dh44R2Mi compared to the control (Figure 7H). Interestingly, under starved conditions, while bmm levels significantly increased in the control, indicating a normal response to starvation, such an increase was not observed in Dh44Mi and Dh44R2Mi. Instead, in these mutants, bmm levels significantly decreased compared to their levels under sated conditions (Figure 7H). This suggests a unique role of DH44 and its receptor Dh44R2 in regulating lipid metabolism and response to nutritional status in Drosophila. To explore the connection between TRPγ and DH44 signaling, we expressed UAS-trpγ under the control of Dh44R1-GAL4 and Dh44R2-GAL4. Only Dh44R2-GAL4 restored the lipid defect in trpγ1, unlike Dh44R1-GAL4 (Figure 7I). Furthermore, Dh44R2-GAL4, but not Dh44R1-GAL4, rescued the excessive lipid accumulation in the FB of trpγ1 mutants (Figure 7J–N). Because Dh44R2 are mostly expressed in the intestine (Figure 7O and P), further study regarding DH44R2 and fat usage are required. Overall, our findings suggest that trpγ expression regulates lipid homeostasis specifically in Dh44 and Dh44R2 cells.

Functional analysis of Dh44, Dh44R1, and Dh44R2 mutants and their roles in lipid accumulation. (A) Total TAG level (µg TAG/mg fly) measurement in whole body extracts from control (w1118), Dh44Mi, Dh44R1Mi, and Dh44R2Mi mutants (n=5). (B–E) Nile red stains of the fat body from w1118, Dh44Mi, Dh44R1Mi, and Dh44R2Mi respectively. Scale bars represent 50 µm. (F) Area of LDs in each indicated genotype (n=3). (G) The protein (µg protein/mg fly) measurement in the whole-body extracts from control (w1118) and Dh44Mi, Dh44R1Mi, and Dh44R2Mi mutants (n=4). (H) Quantification (qRT-PCR) of lipolytic gene (bmm) expression level in the fat body of w1118, Dh44Mi, Dh44R1Mi, and D44R2Mi flies under sated and starved conditions (n=3). (I) Total TAG level measurement in whole body extracts from the indicated genotypes (n=4). (J–M) Nile red stains of the fat body from (J) w1118, (K) trpγ1 (L) trpγ1,UAS-trpγ1/trpγ1,Dh44R1-GAL4, and (M) trpγ1,UAS-trpγ1/trpγ1,Dh44R2-GAL4. (N) Area of LDs in each indicated genotype (n=3). Scale bar represent 50 µm. (O-P) Expression of Dh44R2 in the brain and intestine. (O) Expression of Dh44R2 in the subesophageal zone (SEZ) of the brain. Scale bars represent 50 µm. (P) Expression of Dh44R2 in the intestine. Scale bars represent 200 µm in full intestine and 500 µm in magnified form.

Means ±SEMs. Single factor ANOVA with Scheffe’s analysis was used as a post hoc test to compare multiple sets of data. The asterisks indicate significance from control (*P < 0.05, **P < 0.01). Each dot indicates distribution of individual sample value.

Discussion

Sugars such as D-glucose, D-trehalose, and D-fructose are commonly detected in DH44 neurons in the PI due to the inherent nature and functions of these cells 32. Here, we found that the function of TRPγ, one of the TRPC channels of D. melanogaster, in DH44 neuroendocrine cells plays an essential role in lipid regulation, which is linked to alterations in membrane lipids 33,34. The trpγ1 mutant exhibited clear signs of metabolic syndrome, as demonstrated by reduced carbohydrate level coupled with much higher protein and lipid levels in the body. More importantly, we found that the expression of trpγ was necessary and sufficient to regulate the carbohydrate and lipid levels in the DH44 neurons of trpγ mutants. However, we failed to identify the specific cells required for regulating protein homeostasis because the increased protein levels in the trpγ mutants were not recovered by pilot screening (data not shown) except in the trpγ-GAL4. In conclusion, our findings suggested that the DH44 system has important roles in regulating the metabolic homeostasis of carbohydrates and lipids. Recent studies have suggested that corticotropin-releasing hormone (CRH) is the mammalian homolog of DH44 32. Therefore, we propose that CRH neurons may also regulate lipid levels.

The metabolic dysregulation of carbohydrates and lipids observed in the trpγ mutants was phenocopied by the cellular inactivation of DH44 cells. Compared to the controls, the Dh44-GAL4/UAS-Kir2.1 flies exhibited lower carbohydrate levels and increased lipid levels. We previously proposed that TRPγ holds DH44 neurons in a state of afterdepolarization, thus reducing firing rates by inactivating voltage-gated Na+ channels 11. At the physiological level, this induces the consistent release of DH44 and depletion of DH44 stores, resulting in nutrient utilization and storage malfunctions. Likewise, our findings revealed that Dh44 mutant and its receptor mutant, Dh44R2, displayed defects analogous to those observed in the trpγ mutant. Finally, it would also be interesting to investigate the potential roles of TRPC members in humans, especially the involvement of TRPC4 and TRPC5 in metabolic syndrome.

Multiple lines of evidence have suggested that DH44 neurons communicate with the intestine. First, DH44 at the PI in the brain is an incretin-like hormone that acts as a hemolymph nutrient sensor. The secretion of DH44 is highly dependent on the calorie value of a given diet, as DH44 secretion is only mediated by nutritional sugars but not non-nutritional sugars 32,35. Second, the expression of TRPγ in the PI is sufficient to recover any metabolic defects such as reduced carbohydrate and increased lipid levels, as well as related phenomena such as starvation resistance. DH44 and kinin neuropeptides influence desiccation and starvation tolerance in D. melanogaster 36. Although Dh44Miexhibited normal TAG levels, the presence of larger LDs indicated that DH44 plays a regulatory role in lipid metabolism. LDs consist of a hydrophobic core of neutral storage fats, such as triglycerides or cholesterol esters, encased by a protein-coated phospholipid monolayer. From our findings, it is suggested that DH44 and its receptor DH44R2, but not DH44R1, are involved in the control of lipid storage, indicating a specific pathway through which DH44 influences lipid metabolism distinct from its role in carbohydrate metabolism and overall energy homeostasis. Third, trpγ gene knockdown through RNAi in the DH44 neurons is enough to phenocopy all of the metabolic defects observed in the trpγ mutants. Fourth, trpγ mutants exhibit low expression of brummer lipase, which results in accumulation of LDs in the body, which is a reciprocal process regulated by DH44 neurons. Fifth, the activation of AMPK in the DH44 neurons can recover the defects of lipid level of trpγ1. Sixth, the reduced starvation resistance can be significantly recovered through dietary lipase supplementation. These findings support the assertion that the sugar-sensing DH44 neurons send downward signals to the intestine by absorbing the carbohydrates and lipids derived from dietary intake. However, DH44 neurons do not regulate protein levels in the body because the increased protein levels of trpγ mutants were not recovered by the expression of trpγ in the DH44 neurons. Nevertheless, it is also known that DH44 neurons mediate the influence of dietary amino acids on promoting food intake in flies 37. These findings confirm that trpγ regulates bmm expression in the adult bowel, enabling dietary lipid digestion and maintaining systemic lipid homeostasis.

Treatment with metformin, a lipolytic agent, at two different concentrations (1 mM and 5 mM) decreased the total TAG levels and reduced the size of the LDs in D. melanogaster. Similarly, metformin increased the starvation resistance of the trpγ1 mutants but not the controls. These findings were consistent with the widely acknowledged anti-obesity properties of metformin 21. Furthermore, our results were in agreement with the evidence that metformin helps to reduce lipid accumulation caused by diets rich in saturated fatty acids 38. Although the exact target of metformin is not known, its lipolytic effect is significant and widely documented. Oral administration of 5 mM of metformin reduced normal TAG levels of wild-type flies, whereas the 1 mM dose had no significant effect. Therefore, low metformin doses were effective at decreasing stored lipid levels in flies. The accumulation of intracellular LDs is regulated by autophagy. Metformin improves autophagic flux through AMPKα-mediated mechanisms 38. In the Drosophila context, we propose a mechanism involving AMPK activation within DH44 neurons as a rescue mechanism for starvation resistance in trpγ1, highlighting its crucial role in this process. AMPK activation is known to enhance an organism’s ability to endure environmental stresses by facilitating autophagy-mediated cellular cleanup, promoting mitochondrial biogenesis for increased energy production, and optimizing metabolic pathways for enhanced energy efficiency 38. While the precise nature of the connection between TRPγ, DH44 neuronal activity, and AMPK activation remains unspecified, it is conceivable that TRPγ influences AMPK activation or regulation within DH44 neurons, contributing to stress responses, ATP supply, and metabolic adaptations. Metformin can influence lipid metabolism through various mechanisms, one of which is the promotion of lipolysis 39. This effect results in a gradual reduction of stored lipid levels over time. Orally administering metformin to flies led to a significant reduction in lipid levels, accompanied by an increase in survival. Subsequent qPCR analysis post-metformin feeding revealed a significant upregulation in the expression level of the lipolytic gene (bmm). These results strongly support our hypothesis that metformin induces lipolysis when orally administered.

The exact mechanism through which orally administered lipase leads to gastric lipolysis remains unclear. However, our experimental data indicated that 0.1% lipase increased the survival time of trpγ1 under starvation. Lipase exhibits a high degree of chemical selectivity and has the ability to catalyze triglyceride into glycerol ester, monoglycerides, glycerol, and fatty acids 40. Free fatty acids, HA, and glycerol at concentrations of 0.2%–0.5% can be utilized as energy sources and increase survival time under starvation conditions, suggesting that trpγ1 participates in the utilization of HA and glycerol under starvation conditions.

The therapeutic properties of metformin as a lipolytic and anti-diabetic drug have been widely documented in humans. However, additional studies are needed to identify the exact target of metformin. In this context, our findings demonstrate that D. melanogaster can be an excellent model to study the mode of action and biosafety of metformin.

Materials and methods

Key resources table

Chemical sources

Trehalase from porcine kidney (cat. # 9025-52-9), amyloglucosidase (cat. # 9032-08-0), the glucose (HK) assay kit (cat. # GAHK-20-1KT), and triglyceride mix (cat. # 17810-1AMP-S), lipase from Candida rugosa (cat. # OT705690000), metformin (cat. # D150959), Nile Red (cat. # N3013), glycerol (cat. # G5516) were purchased from Sigma-Aldrich Co. Hexanoic acid (cat. # W255912) were purchased from Wako Pure Chemical Industry Ltd. The PierceTM BCA protein assay kit (cat. # 23225) and the LiquiColor triglyceride test kit (cat. # 2100-225) were purchased from Thermo Fisher Scientific and Stanbio Laboratory, respectively.

TAG level measurement

TAG levels were quantified as described previously with some modifications using a LiquiColor Triglyceride Test kit (cat. # 2100-225, Stanbio Laboratory, Germany) 21. Ten 5–10 day-old male flies were weighed and crushed in 1 mL of PBST (1X PBS and 0.2 percent Triton X-100). The homogenate was incubated at 70 °C for 5 minutes and centrifuged for 3 minutes at 9,500 g. Next, 100 µL of supernatant was transferred into a 1.5 mL Eppendorf tube and mixed with 1 mL of Stanbio LiquiColor Triglyceride Test kit reagent or 1 mL of deionized water to provide a baseline. The reaction mixture was kept at 37 °C for 15 minutes. Finally, the absorbance of the sample solution was measured at 500 nm using a spectrophotometer and the TAG level was calculated based on a standard calibration curve.

Protein level measurement

Protein assays were performed as previously described using the PierceTM BCA Protein Assay Kit with some modifications. Briefly, ten 5–10 day-old male flies were weighed and crushed in 1 mL of PBST (1X PBS and 0.2 percent Triton X-10) and incubated at 70 °C for 5 minutes. The homogenate was then centrifuged for 3 minutes at 9,500 g, after which 300 µL of supernatant was mixed with 600 µL of PierceTM BCA Protein Assay Kit (UF289330). After a 30-minute incubation period at 37 °C, the absorbance of the samples was measured at 530 nm using a spectrophotometer and compared to a standard calibration curve for quantification.

Trehalose and glucose measurements in whole adult flies or hemolymph

Whole-body glucose and trehalose levels were measured in adult flies as previously described 48. Briefly, 10 freshly hatched males were collected, weighed, and crushed in 250 µL of 0.25 M Na2CO3 buffer. The homogenates were then incubated in a water bath (95°C) for 5 min to inactivate all enzymes. Next, 600 µL of 0.25 M sodium acetate and 150 µL of 1 M acetic acid (pH 5.2) were added to the samples, after which the mixtures were centrifuged at 12,500 g for 10 min at 24 °C. Afterward, 200 µL of supernatant was transferred a new microfuge tube and 2 µL porcine kidney trehalase (Sigma: T8778 UN) was added to the sample mixture and incubated overnight at 37 °C to convert trehalose into glucose. Next, 1 mL of glucose hexokinase solution (Sigma: GAHK-20) was added to 100 µL of the sample and incubated for 20 min at 37 °C. Optical density (OD) values were measured at 340 nm. Finally, total glucose and trehalose levels were calculated using a standard glucose curve generated through similar reactions with standard trehalose and glucose.

Glucose and trehalose level measurement in hemolymph were extracted from the male flies. Flies were punctured in the thorax with the help of fine injection needle and kept in 0.5 mL tubes having a punctured base with 21-gauge needle. The punctured flies were adjusted by keeping the shoulder down to prevent leakage from the genital tract. The tubes were set into the 1.5 mL microfuge tube and centrifuged at 4°C for 5 minutes at 2800 g force. 0.5 µL of hemolymph was added to 14.5 µL PBS. 2 µL of porcine kidney trehalase (Sigma: T8778 UN) was added to the sample mixture and incubated overnight at 37°C to convert trehalose into glucose. 1 mL of glucose hexokinase solution (Sigma: GAHK-20) was added to the 100 µL of the sample and incubated at 37°C for 20 minutes. Optical density (OD) values were measured at 340 nm. Total glucose and trehalose levels were calculated using a standard glucose curve generated through similar reactions with standard trehalose and glucose.

Glycogen measurements

Tissue glycogen levels from whole-body extracts of adult flies were quantified as previously described 49. Briefly, 5 males were homogenized in 100 µL of ice-chilled 1X PBS after measuring the weight. The homogenates were kept at 70 °C for 5 min to inactivate all metabolic enzymes. The homogenates were then centrifuged at 12,500 g for 3 min at 4 °C, after which 20 µL of the supernatant was transferred to 1.5 mL microfuge tubes and diluted with 1X PBS to a 1:3 ratio. An amyloglucosidase dilution was prepared by mixing 1.5 µL of amyloglucosidase (Sigma A1602) suspension in 998.5 µL 1X PBS. Finally, a 20 µL aliquot of the diluted sample was added to 20 µL of the diluted amyloglucosidase solution, as well as to the glycogen standard. Both the glycogen standard and test samples were than incubated at 37 °C for 1 h. A commercial glucose (HK) assay reagent (Sigma: G3293 VER) was used to measure total glucose at 340 nm. The glycogen level was quantified by comparing it with standard curve plotted from standard glycogen samples.

Survival assay

Survival experiments were conducted as previously described 50. Normal cornmeal diets were used to conduct the normal survival assays. To measure survival under starvation conditions, twenty 3–4-day-old male flies were fed with 1% agar food supplemented with or without various concentrations of lipase (0.1%; active or denatured), glycerol (1%), hexanoic acid (0.2%, 0.5%), glycerol (1%), TAG mix (0.2%, 0.5%), or metformin (1 mM, 5 mM). Every 12 hours, the flies were monitored/counted and then transferred to fresh vials with the same food supply. The experiments were conducted until the food vials had been cleared.

Immunohistochemistry

We performed immunohistochemistry as previously described 11. Briefly, to fix and block the specimens, we placed freshly dissected tissues (brain and intestine) into a well of 24-well tissue culture plate (Costar Corp.) placed on ice, which contained 940 μL of fixing buffer (1 mM EGTA, 0.1M Pipes pH 6.9, 2 mM MgSO4, 1% TritonX-100, 150 mM NaCl). 60 μL formaldehyde (37%) was add up to the wells and mixed instantly before the tissue were added. As many as tissues that were dissected within 15 minutes were fixed and then the samples were incubated for 30 min more. Samples were washed with wash buffer (1X PBS, 0.1% saponin) for three times (15 min. for each washing), and blocked with 1 mL blocking buffer (1X PBS, 0.1% saponin, and 5 mg/mL BSA) at 4 °C for 4-8 hours.

To perform immunostaining, primary antibodies were inserted into the sample at 4°C for 18 hours 4. Samples were washed with wash buffer three times (15 min each) and added secondary antibodies {(1:200) goat anti-mouse Alexa Fluor 488 (cat. # A11029), goat anti-mouse Alexa Fluor 568 (cat. # A11004), goat anti-rabbit Alexa Fluor 488 (cat. # A11034), and goat anti-rabbit Alexa 568 (cat. # A11036) at 4 °C for 4 hours. Finally, samples were washed three times and stored into 1.25X PDA (187.5 mM NaCl, 37.5% glycerol, 62.5 mM Tris pH 8.8), and kept at 4°C for more than 1 hr. Samples were mounted and examined using a Leica Stellaris 5 Confocal Microscope.

Nile red staining

Nile Red is a dark purplish-red powder (Sigma N-3013), the stock solution must be prepared in acetone (1000 µg/mL) and kept in a tightly sealed, lightproof container at 4 °C. Briefly, 5–10 day-old male flies were fixed in a sagittal position on a glass slide and submerged in a 1 X PBS solution. FBs were gently dissected from a dorsal abdominal region along with thorax or the gut under a stereomicroscope. The dissected tissues were fixed with 4% formaldehyde solution for 15 minutes at room temperature. The fixed tissues were gently washed three times with 1X PBS (5 min for each wash). Nile Red (1:1000 dilution) were added to the tissue samples for 5 minutes. Finally, the stained tissues were washed with 1 mL 1X PBS and mounted in 50% glycerol on a glass slide. LD deposition exhibits greater density in the upper abdominal region compared to the lower abdominal area. This comparison was meticulously conducted within the identical segments 2-3 of the abdomen to ensure accuracy.

For intestinal lipid staining, 5-10 days old male flies were fixed on a glass slide and submerged in 1X PBS solution. Full intestine was dissected very carefully. The dissected intestine was fixed with 4% formaldehyde solution for 15 minutes at room temperature. The fixed intestine was washed 3 times with 1X PBS. Nile red was added to the sample for 5 minutes. The stained samples were washed with 1mL 1X PBS and mounted in 50% glycerol on a glass slide. The stained tissues were examined using a Leica Stellaris 5 Confocal Microscope.

qRT-PCR

qRT-PCR assays were performed as described previously 51. Briefly, ten 6–10 day-old male flies (control and mutant) were selected for whole body and 15 flies were selected for fat body samples in Supplement Figure 6K. The experiments were conducted under sated (0 h starvation) and starved (24 h starvation) conditions. Total RNA was extracted using the TRizol reagent (Invitrogen) followed by DNase (Promega) treatment. cDNA was synthesized using the AMV reverse transcriptase system (Promega). RT-qPCR experiments were carried out using a Bio-rad CFX system. The Takara TB Green Premix was used to assess the mRNA expression level of each gene according to the manufacturer’s instructions. Relative gene expression was calculated using the 2−ΔΔCt method. Three biological samples were used and transcript levels were normalized to the D. melanogaster housekeeping gene tubulin. The experiments were conducted using the following primer pairs: acc, 5’-ACG AGG GCG AGC AGC GTT AC-3′ (forward) and 5′-TAG GGC GAC TTG GTG GGC AT-3′ (reverse); bmm, 5′-ATG ACT TCG GAC TTC TTC AGG G-3′ (forward) and 5′-CCA ATT CAG ATG GAA GAG CTG-3′ (reverse); fbp, 5′-CTC CAA CGA GCT GTT CAT CA-3′ (forward) and 5′-TGA ACC GAT CGA CAC CAG GC-3′ (reverse); pepck1, 5′-AGG TGC ACA TCT GCG ATG GC-3′ (forward) and 5′-CCA CCA CGT AAG CAG AGT CC-3′ (reverse); desat1, 5’-AAG CCG GTG CCC AGT CCA TC-3’ (forward) and 5’-ATG GTC GCG AGC CCA ATG GT-3’ (reverse); and tubulin, 5′-TCC TTG TCG CGT GTG AAA CA-3′ (forward) and 5′-CCG AAC GAG TGG AAG ATG AG-3’ (reverse).

Statistics and Reproducibility

D. melanogaster was selected as a model organism in this study. For the experiments, male flies were mostly used unless we did not mention the sex. All of the experiments were conducted under laboratory conditions. The appropriate number of replicates was established based on previous research. A large enough sample size was used in all of our assays to ensure that our results were representative and repeatable. No data points were left out of the analysis. For each genotype, the data points indicate the values of individual replicates. The error bars in all of the figures represent the standard error of the mean (SEM). The analysis of the RT-qPCR data was conducted using the CT values. Comparisons between multiple experimental groups were conducted via single-factor ANOVA and Scheffe’s post hoc test. Pair-wise comparisons were conducted via Student’s t-test. Survival curves were estimated for each group, using a Kaplan-Meier method and compared statistically using the log-rank tests. The asterisks in the figures indicate statistical significance (*P<0.05, **P<0.01). All statistical analyses were conducted using the Origin Pro 8 software for Windows (ver. 8.0932; Origin Lab Corporation, USA).

eLife Assessment

This important study using Drosophila genetics explores the role of TRPγ in Dh44 neuroendocrine cells for lipid and protein metabolism. Evidence for lipid storage and metabolism measured by triacylglycerol levels, lipid droplet size, and starvation resistance are generally solid to support the conclusion. However, the claim on the TRPγ functions in Dh44R2 is still unclear, as the analysis of the role and expression of Dh44R2 in the gut is incomplete.

Reviewer #1 (Public Review):

Summary:

This research article by Nath et al. from the Lee Lab addresses how lipolysis under starvation is achieved by a transient receptor potential channel, TRPγ, in the neuroendocrine neurons to help animals survive prolonged starvation. Through a series of genetic analyses, the authors identify that TRPγ mutations specifically lead to a failure in lipolytic processes under starvation, thereby reducing animals' starvation resistance. The conclusion was confirmed through total triacylglycerol levels in the animals and lipid droplet staining in the fat bodies. This study highlights the importance of transient receptor potential (TRP) channels in the fly brain to modulate energy homeostasis and combat metabolic stress. While the data is compelling and the message is easy to follow, several aspects require further clarification to improve the interpretation of the research and its visibility in the field.

Strengths:

This study identifies the biological meaning of TRPγ in promoting lipolysis during starvation, advancing our knowledge about TRP channels and the neural mechanisms to combat metabolic stress. Furthermore, this study demonstrates the potential of the TRP channel as a target to develop new therapeutic strategies for human metabolic disorders by showing that metformin and AMPK pathways are involved in its function in lipid metabolisms during starvation in Drosophila.

Weaknesses:

Some key results that might strengthen their conclusions were left out for discussion or careful explanation (see below). If the authors could improve the writing to address their findings and connect their findings with conclusions, the research would be much more appreciated and have a higher impact in the field.

Here, I listed the major issues and suggestions for the authors to improve their manuscript:

(1) Are the increased lipid droplet size and the upregulated total TAG level measured in the starved or sated mutant in Figure 1? This information might be crucial for readers to understand the physiological function of TRP in lipid metabolism. In other words, clarifying whether the upregulated lipid storage is observed only in the starved trp mutant will advance our knowledge of TRPγ. If the increase of total TAG level is only observed in the starved animals, TRP in the Dh44 neurons might serve as a sensor for the starvation state required to promote lipolysis in starvation conditions. On the other hand, if the total TAG level increases in both starved and sated animals, activation of Dh44 through TRPγ might be involved in the lipid metabolism process after food ingestion.

(2) It is unclear how AMPK activation in Dh44 neurons reduces the total triacylglycerol (TAG) levels in the animals (Figure 3G). As AMPK is activated in response to metabolic stress, the result in Figure 3G might suggest that Dh44 neurons sense metabolic stress through AMPK activation to promote lipolysis in other tissues. Do Dh44 neurons become more active during starvation? Is activation of Dh44 neurons sufficient to activate AMPK in the Dh44 neurons without starvation? Is activation of AMPK in the Dh44 neurons required for Dh44 release and lipolysis during starvation? These answers would provide more insights into the conclusion in Lines 192-193.

(3) It is unclear how the lipolytic gene brummer is further downregulated in the trpγ mutant during starvation while brummer is upregulated in the control group (Figure 6A). This result implies that the trpγ mutant was able to sense the starvation state but responded abnormally by inhibiting the lipolytic process rather than promoting lipolysis, which makes it more susceptible to starvation (Figure 3B).

(4) There is an inconsistency of total TAG levels and the lipid droplet size observed in the Dh44 mutant but not in the Dh44-R2 mutant (Figures 7A and 7F). This inconsistency raises a possibility that the signaling pathway from Dh44 release to its receptor Dh44-R2 only accounts for part of the lipid metabolic process under starvation. Adding discussion to address this inconsistency may be helpful for readers to appreciate the finding.

Reviewer #2 (Public Review):

Summary:

In this paper, the function of trpγ in lipid metabolism was investigated. The authors found that lipid accumulation levels were increased in trpγ mutants and remained high during starvation; the increased TAG levels in trpγ mutants were restored by the expression of active AMPK in DH44 neurons and oral administration of the anti-diabetic drug metformin. Furthermore, oral administration of lipase, TAG, and free fatty acids effectively restored the survival of trpγ mutants under starvation conditions. These results indicate that TRPv plays an important role in the maintenance of systemic lipid levels through the proper expression of lipase. Furthermore, authors have shown that this function is mediated by DH44R2. This study provides an interesting finding in that the neuropeptide DH44 released from the brain regulates lipid metabolism through a brain-gut axis, acting on the receptor DH44R2 presumably expressed in gut cells.

Strengths:

Using Drosophila genetics, careful analysis of which cells express trpγ regulates lipid metabolism is performed in this study. The study supports its conclusions from various angles, including not only TAG levels, but also fat droplet staining and survival rate under starved conditions, and oral administration of substances involved in lipid metabolism.

Weaknesses:

Lipid metabolism in the gut of DH44R2-expressing cells should be investigated for a better understanding of the mechanism. Fat accumulation in the gut is not mechanistically linked with fat accumulation in the fat body. The function of lipase in the gut (esp. R2 region) should be addressed, e.g. by manipulating gut-lipases such as magro or Lip3 in the gut in the contest of trpγ mutant. Also, it is not clarified which cell types in the gut DH44R2 is expressed. The study also mentioned only in the text that bmm expression in the gut cannot restore lipid droplet enlargement in the fat body, but this result might be presented as a figure.

Reviewer #3 (Public Review):

In this manuscript, the authors demonstrated the significance of the TRPγ channel in regulating internal TAG levels. They found high TAG levels in TRPγ mutant, which was ascribed to a deficit in the lipolysis process due to the downregulation of brummer (bmm). It was notable that the expression of TRPγ in DH44+ PI neurons, but not dILP2+ neurons, in the brain restored the internal TAG levels and that the knockdown of TRPγ in DH44+ PI neurons resulted in an increase in TAG levels. These results suggested a non-cell autonomous effect of Dh44+PI neurons. Additionally, the expression of the TRPγ channel in Dh44 R2-expressing cells restored the internal TAG levels. The authors, however, did not provide an explanation of how TRPγ might function in both presynaptic and postsynaptic cells in the non-cell autonomous manner to regulate the TAG storage. The authors further determined the effect of TRPγ mutation on the size of lipid droplets (LD) and the lifespan and found that TRPγ mutation caused an increase in the size of LD and a decrease in the lifespan, which were reverted by feeding lipase and metformin. These were creative endeavors, I thought. The finding that DH44+ PI neurons have non-cell autonomous functions in regulating bodily metabolism (mainly sugar/lipid) in addition to directing sugar nutrient sensing and consumption is likely correct, but the paper has many loose ends. I would like to see a revision that includes more experiments to tighten up the findings and appropriate interpretations of the results.

(1) The authors need to provide interpretations or speculations as to how DH44+ PI neurons have non-cell autonomous functions in regulating the internal TAG stores, and how both presynaptic DH44 neurons and postsynaptic DH44 R2 neurons require TRPγ for lipid homeostasis.

(2) The expression of TRPγ solely in DH44 R2 neurons of TRPγ mutant flies restored the TAG phenotype, suggesting an important function mediated by TRPγ in DH44 R2 neurons. However, the authors did not document the endogenous expression of TRPγ in the DH44R2+ gut cells. This needs to be shown.

(3) While Dh44 mutant flies displayed normal internal TAG levels, Dh44R2 mutant flies exhibited elevated TAG levels (Figure 7A). This suggested that the lipolysis phenotype could be facilitated by a neuropeptide other than Dh44. Alternatively, a Dh44 neuropeptide-independent pathway could mediate the lipolysis. In either case, an additional result is needed to substantiate either one of the hypotheses.

(4) While the authors observed an increased area of fat body lipid droplets (LD) in Dh44 mutant flies (Figure 7F), they did not specify the particular region of the fat body chosen for measuring the LD area.

(5) The LD area only accounts for TAG levels in the fat body, whereas TAG can be found in many other body parts, including the R2 area as demonstrated in Figure 5A-D using Nile red staining. As such, measuring the total internal TAG levels would provide a more accurate representation of TAG levels than the average fat body LD area.

(6) In Figure 5F-I, the authors should perform the similar experiment with Dh44, Dh44R1, and Dh44R2 mutant flies.

(7) The representative image in Figure 6B does not correspond to the GFP quantification results shown in Figure 6C. In trpr1;bmm::GFP flies, the GFP signal appears stronger in starved conditions than in satiated conditions.

(8) In Figure 6H-I, fat body-specific expression of bmm reversed the increased LD area in TRPγ mutants. The authors also showed that Dh44+PI neuron-specific expression of bmm yielded a similar result. The authors need to provide an interpretation as to how bmm acts in the fat body or DH44 neurons to regulate this.

(9) The authors should explain why the DH44 R1 mutant did not represent similar results as the wild type.

(10) It would be good to have a schematic that represents the working model proposed in this manuscript.

Author response:

The following is the authors’ response to the original reviews

Public Reviews:

Reviewer #1 (Public Review):

Summary:

This research article by Nath et al. from the Lee Lab addresses how lipolysis under starvation is achieved by a transient receptor potential channel, TRPγ, in the neuroendocrine neurons to help animals survive prolonged starvation. Through a series of genetic analyses, the authors identify that TRPγ mutations specifically lead to a failure in lipolytic processes under starvation, thereby reducing animals' starvation resistance. The conclusion was confirmed through total triacylglycerol levels in the animals and lipid droplet staining in the fat bodies. This study highlights the importance of transient receptor potential (TRP) channels in the fly brain to modulate energy homeostasis and combat metabolic stress. While the data is compelling and the message is easy to follow, several aspects require further clarification to improve the interpretation of the research and its visibility in the field.

Strengths:

This study identifies the biological meaning of TRPγ in promoting lipolysis during starvation, advancing our knowledge about TRP channels and the neural mechanisms to combat metabolic stress. Furthermore, this study demonstrates the potential of the TRP channel as a target to develop new therapeutic strategies for human metabolic disorders by showing that metformin and AMPK pathways are involved in its function in lipid metabolisms during starvation in Drosophila.

Weaknesses:

Some key results that might strengthen their conclusions were left out for discussion or careful explanation (see below). If the authors could improve the writing to address their findings and connect their findings with conclusions, the research would be much more appreciated and have a higher impact in the field.

Here, I listed the major issues and suggestions for the authors to improve their manuscript:

(1) Are the increased lipid droplet size and the upregulated total TAG level measured in the starved or sated mutant in Figure 1? This information might be crucial for readers to understand the physiological function of TRP in lipid metabolism. In other words, clarifying whether the upregulated lipid storage is observed only in the starved trp mutant will advance our knowledge of TRPγ. If the increase of total TAG level is only observed in the starved animals, TRP in the Dh44 neurons might serve as a sensor for the starvation state required to promote lipolysis in starvation conditions. On the other hand, if the total TAG level increases in both starved and sated animals, activation of Dh44 through TRPγ might be involved in the lipid metabolism process after food ingestion.

We measured total TAG level in Figure 1 and LD sizes in Figure 2 under sated condition. We inserted “under sated condition” to clarify it. lines 97 and 147-148.

Thanks for your suggestions.

(2) It is unclear how AMPK activation in Dh44 neurons reduces the total triacylglycerol (TAG) levels in the animals (Figure 3G). As AMPK is activated in response to metabolic stress, the result in Figure 3G might suggest that Dh44 neurons sense metabolic stress through AMPK activation to promote lipolysis in other tissues. Do Dh44 neurons become more active during starvation? Is activation of Dh44 neurons sufficient to activate AMPK in the Dh44 neurons without starvation? Is activation of AMPK in the Dh44 neurons required for Dh44 release and lipolysis during starvation? These answers would provide more insights into the conclusion in Lines 192-193.

In our previous study, we demonstrated that trpγ mutants exhibited lower levels of glucose, trehalose and glycogen level (Dhakal et al. 2022), and in the current study, we observed excessive lipid storage in the trpγ mutant, indicating imbalanced energy homeostasis. Given the established role of AMPK in maintaining energy balance (Marzano et. al., 2021, Lin et al 2021), we employed the activated form of AMPK (UAS-AMPKTD) in our experiments. Our result showed that expression of activated AMPK in Dh44 neurons led to a reduction in total TAG levels, suggesting that AMPK activation in these neurons can promote lipolysis even in the absence of starvation. Regarding the activation of Dh44 neurons, Dus et al in 2015 reported that Dh44 cells in the brain are activated by nutritive sugars especially in starvation conditions. In addition, another report showed a role of Dh44 neuron in regulating starvation induced sleep suppression (Oh et. al., 2023) which may imply that these neurons become more active under starved conditions. We did not directly assess whether Dh44 neuron activity increases during starvation or whether AMPK activation in these neurons is required for DH44 release and subsequent lipolysis, our finding support the notion that AMPK activation in Dh44 neuron is sufficient to reduce TAG levels, potentially by metabolic stress response typically observed during starvation. We explained it like the following: “Dh44 neurons regulate starvation-induced sleep suppression (Oh et. al., 2023), which implies that these neurons become more active under starved conditions.” lines 190-191.

(3) It is unclear how the lipolytic gene brummer is further downregulated in the trpγ mutant during starvation while brummer is upregulated in the control group (Figure 6A). This result implies that the trpγ mutant was able to sense the starvation state but responded abnormally by inhibiting the lipolytic process rather than promoting lipolysis, which makes it more susceptible to starvation (Figure 3B).

Thanks for your suggestions. We explained it like the following: “The data indicates that the trpg mutant can sense the starvation state but responds abnormally by suppressing lipolysis instead of activating it. This dysregulated lipolytic response likely increases the mutant's vulnerability to starvation, as it cannot effectively mobilize lipid stores for energy during periods of nutrient deprivation.” lines 251-254.

(4) There is an inconsistency of total TAG levels and the lipid droplet size observed in the Dh44 mutant but not in the Dh44-R2 mutant (Figures 7A and 7F). This inconsistency raises a possibility that the signaling pathway from Dh44 release to its receptor Dh44-R2 only accounts for part of the lipid metabolic process under starvation. Adding discussion to address this inconsistency may be helpful for readers to appreciate the finding.

Thanks for your suggestion. We included the following in the Discussion: “There is an inconsistency of total TAG levels and the LD size observed in the Dh44 mutant. This inconsistency raises a possibility that the signaling pathway from DH44 release to its receptor DH44R2 only accounts for part of the lipid metabolic process under starvation. While Dh44 mutant flies displayed normal internal TAG levels, Dh44R2 mutant flies exhibited elevated TAG levels. This suggested that the lipolysis phenotype could be facilitated by a neuropeptide other than DH44. Alternatively, a DH44 neuropeptide-independent pathway could mediate the lipolysis.” lines 429-436.

Reviewer #2 (Public Review):

Summary:

In this paper, the function of trpγ in lipid metabolism was investigated. The authors found that lipid accumulation levels were increased in trpγ mutants and remained high during starvation; the increased TAG levels in trpγ mutants were restored by the expression of active AMPK in DH44 neurons and oral administration of the anti-diabetic drug metformin. Furthermore, oral administration of lipase, TAG, and free fatty acids effectively restored the survival of trpγ mutants under starvation conditions. These results indicate that TRPv plays an important role in the maintenance of systemic lipid levels through the proper expression of lipase. Furthermore, authors have shown that this function is mediated by DH44R2. This study provides an interesting finding in that the neuropeptide DH44 released from the brain regulates lipid metabolism through a brain-gut axis, acting on the receptor DH44R2 presumably expressed in gut cells.

Strengths:

Using Drosophila genetics, careful analysis of which cells express trpγ regulates lipid metabolism is performed in this study. The study supports its conclusions from various angles, including not only TAG levels, but also fat droplet staining and survival rate under starved conditions, and oral administration of substances involved in lipid metabolism.

Weaknesses:

Lipid metabolism in the gut of DH44R2-expressing cells should be investigated for a better understanding of the mechanism. Fat accumulation in the gut is not mechanistically linked with fat accumulation in the fat body. The function of lipase in the gut (esp. R2 region) should be addressed, e.g. by manipulating gut-lipases such as magro or Lip3 in the gut in the contest of trpγ mutant. Also, it is not clarified which cell types in the gut DH44R2 is expressed. The study also mentioned only in the text that bmm expression in the gut cannot restore lipid droplet enlargement in the fat body, but this result might be presented as a figure.

We appreciate the reviewer’s insightful suggestions. Unfortunately, due to the unviability of the reagent (UAS-Lip3), we were unable to manipulate gut lipase in trpy mutants as proposed. However, we additionally performed immunostaining to examine the co-expression of trpγ and Dh44R2 in the gut, and our results indicate that both trpγ and Dh44R2 are co-expressed in the R2 region of the gut (Figure 7O and P). Furthermore, we have updated our figures to address the point that bmm expression in the gut does not restore lipid droplet enlargement in the fat body, with the revised version (Figure 5I and J).

Reviewer #3 (Public Review):

In this manuscript, the authors demonstrated the significance of the TRPγ channel in regulating internal TAG levels. They found high TAG levels in TRPγ mutant, which was ascribed to a deficit in the lipolysis process due to the downregulation of brummer (bmm). It was notable that the expression of TRPγ in DH44+ PI neurons, but not dILP2+ neurons, in the brain restored the internal TAG levels and that the knockdown of TRPγ in DH44+ PI neurons resulted in an increase in TAG levels. These results suggested a non-cell autonomous effect of Dh44+PI neurons. Additionally, the expression of the TRPγ channel in Dh44 R2-expressing cells restored the internal TAG levels. The authors, however, did not provide an explanation of how TRPγ might function in both presynaptic and postsynaptic cells in the non-cell autonomous manner to regulate the TAG storage. The authors further determined the effect of TRPγ mutation on the size of lipid droplets (LD) and the lifespan and found that TRPγ mutation caused an increase in the size of LD and a decrease in the lifespan, which were reverted by feeding lipase and metformin. These were creative endeavors, I thought. The finding that DH44+ PI neurons have non-cell autonomous functions in regulating bodily metabolism (mainly sugar/lipid) in addition to directing sugar nutrient sensing and consumption is likely correct, but the paper has many loose ends. I would like to see a revision that includes more experiments to tighten up the findings and appropriate interpretations of the results.

(1) The authors need to provide interpretations or speculations as to how DH44+ PI neurons have non-cell autonomous functions in regulating the internal TAG stores, and how both presynaptic DH44 neurons and postsynaptic DH44 R2 neurons require TRPγ for lipid homeostasis.

In Discussion, we had mentioned our previous finding. “ We previously proposed that TRPg holds DH44 neurons in a state of afterdepolarization, thus reducing firing rates by inactivating voltage-gated Na+ channels (Dhakal et al., 2022). At the physiological level, this induces the consistent release of DH44 and depletion of DH44 stores, resulting in nutrient utilization and storage malfunctions.”

We also included the following: “TRPg in DH44 neurons may influence the release of metabolic signals or hormones that act on postsynaptic DH44R2 cells. These postsynaptic cells could, in turn, modulate lipid storage and metabolism in a non-cell autonomous manner. However, the mechanism by which TRPg functions in DH44R2 cells remains unclear. One possible explanation is that TRPg in the gut may be activated by stretch or osmolarity (Akitake et al. 2015).” lines 439-440.

This interaction between presynaptic and postsynaptic cells may ensure a coordinated response to metabolic changes and maintain lipid homeostasis. Thus, both Dh44-expressing and Dh44-R2-expressing cells are crucial for the proper functioning of TRPγ in regulating internal TAG levels and lipid storage.

(2) The expression of TRPγ solely in DH44 R2 neurons of TRPγ mutant flies restored the TAG phenotype, suggesting an important function mediated by TRPγ in DH44 R2 neurons. However, the authors did not document the endogenous expression of TRPγ in the DH44R2+ gut cells. This needs to be shown.

We appreciate the reviewer’s suggestion. To address this, we performed immunostaining to examine the expression of TRPγ in the DH44R2+ gut cells. Our results, as shown in Figure 7 O and P, confirm that TRPγ is co-expressed in the Dh44R2+ cells in the gut. We also found that Dh44R2 is expressed in the brain as well. We documented this part like the following: “Given that Dh44R2 is predominantly expressed in the intestine, we performed immunostaining to examine whether Dh44R2 co-localizes with trpg in gut cells. Our results confirmed that Dh44R2 and trpg are co-expressed in intestinal cells (Figure 7O and P). Additionally, we analyzed Dh44R2 expression in the brain and found that two Dh44R2-expressing cells are co-localized with Dh44-expressing cells in the PI region (Figure 7Q). To further delineate whether Dh44R2-mediated fat utilization is specific to the brain, gut, or fat body, we knocked down Dh44R2RNAi using Dh44-GAL4, myo1A-GAL4, and cg-GAL4, respectively (Figure 7–figure supplement 1E). Notably, knockdown of Dh44R2 with Myo1A-GAL4 resulted in elevated TAG levels, indicating that DH44R2 activity in lipid metabolism is specific to the gut.” lines 375-384.

(3) While Dh44 mutant flies displayed normal internal TAG levels, Dh44R2 mutant flies exhibited elevated TAG levels (Figure 7A). This suggested that the lipolysis phenotype could be facilitated by a neuropeptide other than Dh44. Alternatively, a Dh44 neuropeptide-independent pathway could mediate the lipolysis. In either case, an additional result is needed to substantiate either one of the hypotheses.

The Dh44 mutant flies exhibited normal TAG levels, whereas Dh44R2 mutant flies showed elevated TAG levels. However, when we examined the lipid droplets in the fat body, both Dh44 mutant and Dh44R2 mutant flies displayed larger lipid droplets, indicating a disruption in lipid metabolism. Additionally, we assessed starvation survival time and found that both Dh44 and Dh44R2 mutant flies exhibited reduced survival under starvation conditions compared to controls. Supplementation with lipase (Figure 7–figure supplement 1A), glycerol (Figure 7–figure supplement 1B), hexanoic acid (Figure 7–figure supplement 1C), and mixed TAGs (Figure 7–figure supplement 1D) improved starvation survival time, further supporting that the lipid metabolism pathway was impaired in both mutants. These observations highlight the role of Dh44 in regulating lipolysis. We included related Discussion: “There is an inconsistency of total TAG levels and the LD size observed in the Dh44 mutant. This inconsistency raises a possibility that the signaling pathway from DH44 release to its receptor DH44R2 only accounts for part of the lipid metabolic process under starvation. While Dh44 mutant flies displayed normal internal TAG levels, Dh44R2 mutant flies exhibited elevated TAG levels. This suggested that the lipolysis phenotype could be facilitated by a neuropeptide other than DH44. Alternatively, a DH44 neuropeptide-independent pathway could mediate the lipolysis.” lines 429-436.

(4) While the authors observed an increased area of fat body lipid droplets (LD) in Dh44 mutant flies (Figure 7F), they did not specify the particular region of the fat body chosen for measuring the LD area.

We have chosen the 2-3 segment in the abdomen for all fat body images, which we already mentioned in Nile red staining in the Method section line 630-631.

(5) The LD area only accounts for TAG levels in the fat body, whereas TAG can be found in many other body parts, including the R2 area as demonstrated in Figure 5A-D using Nile red staining. As such, measuring the total internal TAG levels would provide a more accurate representation of TAG levels than the average fat body LD area.

We have measured total internal TAG level in whole body throughout the experiments (Figure 1F, 2C, 2E, 3C, 3G, 4A, 4B, 7A, 7I, and many Supplementary Figures) except bmm expression using GAL4/UAS system. Now we include this new data in Figure 5–figure supplement 1) which is the same conclusion with LD analysis.

(6) In Figure 5F-I, the authors should perform the similar experiment with Dh44, Dh44R1, and Dh44R2 mutant flies.

We did the experiments with Dh44, Dh44R1, and Dh44R2 mutant flies and we found that Dh44 and Dh44R2 mutant flies showed reduced starvation survival time than control and which was increased after supplementation of lipase, glycerol, hexanoic acid and TAG (Figure 7– figure supplement 1A–D). lines 361-372.

(7) The representative image in Figure 6B does not correspond to the GFP quantification results shown in Figure 6C. In trpr1;bmm::GFP flies, the GFP signal appears stronger in starved conditions than in satiated conditions.

We updated it with new images. We quantified GFP intensity level using image J and found that GFP intensity level was significantly lower in starved condition in trpγ1;bmm::GFP flies than sated condition.

(8) In Figure 6H-I, fat body-specific expression of bmm reversed the increased LD area in TRPγ mutants. The authors also showed that Dh44+PI neuron-specific expression of bmm yielded a similar result. The authors need to provide an interpretation as to how bmm acts in the fat body or DH44 neurons to regulate this.

We first inserted the following in results: “Furthermore, the expression of bmm in the fat body, as well as Dh44 neurons in the PI region, can promote lipolysis at the systemic level.” lines 276-277.

Additionally, we discussed it in the Discussion: “Brummer lipase is essential for regulating lipid levels in the insect fat body by mediating lipid mobilization and energy homeostasis. In Nilaparvata lugens, it facilitates triglyceride breakdown (Lu et al., 2018), while studies in Drosophila show that reduced Brummer lipase expression decreases fatty acids and increases diacylglycerol levels, highlighting its role in lipid metabolism (Nazario-Yepiz et al., 2021). Here, we additionally demonstrate that bmm expression in DH44 neurons within the PI region can systemically regulate TAG levels. Cell signaling or energy status in DH44 neurons may contribute to hormonal release that targets organs such as the fat body.” lines 451-459.

(9) The authors should explain why the DH44 R1 mutant did not represent similar results as the wild type.

We added “In addition, bmm levels in Dh44R1Mi under starved condition did not increase as significantly as in the control. This suggests a unique role of DH44 and its receptors in regulating lipid metabolism and response to nutritional status in Drosophila.” lines 358-360.

(10) It would be good to have a schematic that represents the working model proposed in this manuscript.

We updated the schematic model in revised version (Figure 8).

Recommendations for the authors:

Reviewing Editor (Recommendations For The Authors):

This paper characterized the function of trpγ in Dh44-expressing PI neurons for lipid metabolism and lipolysis induced by prolonged starvation. The authors applied a series of lipolytic genetic manipulation and lipid/lipid metabolism supplements to rescue the trpγ deficits in lipolysis: the expression of active AMPK in the DH44-expressing PI neurons or brummer, a lipolytic gene, in the trpγ-expressing cells, and oral administration of the anti-diabetic drug metformin, lipase, TAG and free fatty acids. Despite this exhaustive characterization of the defective lipolysis in the trpγ mutants, there remain puzzles in inconsistent defects of Dh44 and DH44R2 in the total TAG levels and in the expression and functions of the receptor in the gut. Clarification of these points and other issues raised by the reviewers should improve the mechanisms of lipid metabolism through Dh44 signalling.

Reviewer #1 (Recommendations For The Authors):

(1) It might be worth introducing Dh44 in the introduction section as it is unclear to readers how the authors hypothesized the site-of-action of TRPγ in Dh44 neurons for lipid metabolism after reading the introduction.

We introduced the following: “We found that TRPg expression in Dh44 neuroendocrine cells in the brain is critical for maintaining normal carbohydrate levels in tissues (Dhakal et al. 2022). Building on this, we hypothesized that TRPg in Dh44 cells also regulates lipid and protein homeostasis.” lines 69-71.

(2) Providing a summary model in the end to integrate the present findings and their previous publication about TRPγ functions in Drosophila sugar selection would greatly help readers understand and appreciate the general role of TRPγ in balancing energy homeostasis.

We made a schematic model in Figure 8.

(3) Swapping the order of Figures 5 and 6 might be a better way to tell the story without logic gaps. The results addressing the mechanisms of metformin and TRPγ in promoting lipolysis under starvation are interrupted by the lipid storage data in the R2 cells in the current Figure 5A-5E. In addition, presenting Figure 5A-5E before or together with Figure 7 will help readers appreciate the expression of Dh44-R2 and its function in regulating lipid metabolism in Figure 7.

We did.

(4) It might be misleading to use the word "sated" for the condition of 5-hour mild starvation. The word "mild starvation" or the equivalents might be a better word choice.

We appreciate the reviewer’s concern. As hemolymph sugar level does not drop down significantly in 5 hr starvation, the previous papers (Dus et al 2015, Dhakal et al 2022) indicated it as sated condition. To use the word consistently, we prefer using “sated” instead of “mild starvation”.

(5) It is unclear what the white arrows are pointing at in Figures 7O and 7P. Some of those seem to be non-specific signals, so it is hard to connect the figure to the conclusion in Lines 351-353. It would be helpful to add some explanations to help readers interpret Figures 7O and 7P.

In the previous version, Figure 7O and 7P white arrows represented the expression of Dh44R2 in the SEZ region of the brain and R2 region of the gut. In revised version, to make clear, we performed additional immunostaining for the co-expression of trpγ and Dh44R2 in the gut. We found that trpγ and Dh44R2 co-expressed at the R2 region of the gut specifically (Figure 7O and P). Similarly, we found that two cells of Dh44R2 co-expressed in Dh44 cells in the PI region of the brain (now Figure 7Q). We updated this part. lines 375-380.

(6) The figure legend for the (G) panel in Figure 2-figure Supplement 1 was mislabeled as (F).

We corrected it.

(7) In Line 85, the authors might want to write "… among these mutants, only trpγ mutant displayed reduced carbohydrate levels, suggesting …". Please confirm the information for the sentence. lines 87-88.

We clarified it.

Reviewer #2 (Recommendations For The Authors):

(1) The trpγ[G4] would be difficult for non-Drosophila researchers to understand; it would be better to use trpγ-Gal4.

We got the mutant line from Dr. Craig Montell who named it. We explained it like the following in the main text: “controlled by GAL4 knocked into the trpg locus (trpgG4 flies; +)” line 109.

(2) The arrows in Figures 7O and 7P need to be explained in the figure legends.

We did.

Reviewer #3 (Recommendations For The Authors):

(11) Lines 95-96 should have a reference.

We did.

(12) Lines 129-130: It should read "TRPγ expressed in DH44 cells is sufficient for the regulation of lipid levels."

We changed it as suggested.

(13) Figure 5E needs to be repeated with more trials.

We increased the n numbers. Previously (Figure 5E) we included area of 10 LDs from 3 samples, and in revised figure (Figure 6I) we have included 28 LDs from 10 samples.

(14) Figures 5F-I, bold lines are not too visible and therefore, dotted lines could be used.

We changed it as suggested.

(15) Line 356: It is not true that D-trehalose or D-fructose is commonly detected by DH44 neurons. These sugars at concentrations much higher than the physiological concentration range stimulate DH44 neurons (see Dus et al., 2015).

We removed it.

(16) Lines 362-363: It should read "Expression of TRPγ in DH44 neurons was necessary and sufficient to regulate the carbohydrate and lipid levels.".

We changed it.

(17) Lines 369-370: The authors need to consider removing the possible role of CRF in regulating lipid homeostasis. It could be considered to be far-fetched.

We removed it.

(18) Line 407-408: the sentence "Nevertheless, it is also known that DH44 neurons mediate the influence of dietary amino acids on promoting food intakes in flies (37)" needs to be removed. They used amino acid concentrations that were far greater than the physiological levels observed in the internal milieu of flies. Still, many laboratories cannot reproduce the result of using the high AA concentrations.

We removed it.

Version 2

Abstract

Abstract

Understanding how the brain controls nutrient storage is pivotal. Transient receptor potential (TRP) channels are conserved from insects to humans. They serve in detecting environmental shifts and in acting as internal sensors. Here, we found that a TRPγ mutant exhibited in Drosophila melanogaster are required for maintaining normal lipid and protein levels. In animals, lipogenesis and lipolysis control lipid levels in response to food availability. Lipids are mostly stored as triacylglycerol in the fat bodies (FBs) of D. melanogaster. Interestingly, trpγ deficient mutants exhibited elevated TAG levels and our genetic data indicated that Dh44 neurons are indispensable for normal lipid storage but not protein storage. The trpγ mutants also exhibited reduced starvation resistance, which was attributed to insufficient lipolysis in the FBs. This could be mitigated by administering lipase or metformin orally, indicating a potential treatment pathway. Gene expression analysis indicated that trpγ knockout downregulated brummer, a key lipolytic gene, resulting in chronic lipolytic deficits in the gut and other fat tissues. The study also highlighted the role of specific proteins, including neuropeptide DH44 and its receptor DH44R2 in lipid regulation. Our findings provide insight into the broader question of how the brain and gut regulates nutrient storage.

Introduction

Nutrient storage is not a simple matter of accumulation but involves a sophisticated network of signals and receptors, feedback loops, and hormonal influences, all choreographed by the brain. The mechanisms involve various tissues and organs, including the liver, muscles, and adipose tissue, each playing a distinct role in the storage and release of vital nutrients like glucose, fats, and proteins in mammals (Efeyan et al., 2015). Key to this regulation are specialized cells and molecular pathways that respond to dietary intake, energy expenditure, and changes in the internal and external environment (Tran et al., 2022). For instance, the brain receives and interprets signals about the body’s energy status through hormones such as insulin and leptin, and nutrients themselves, which inform decisions on whether to store or mobilize energy reserves in mammals (Jais and Brüning, 2022; Roh and Kim, 2016). Moreover, the role of Transient Receptor Potential (TRP) channels as molecular sensors that detect changes in the environment and internal metabolic status highlights the level of molecular sophistication involved in nutrient storage and energy balance (Clapham, 2003; Dhakal and Lee, 2019; Fowler and Montell, 2013; Turner et al., 2016). These channels bridge external stimuli with internal responses, ensuring the body adapts efficiently to maintain homeostasis (Dhakal and Lee, 2019; Venkatachalam and Montell, 2007). This concert of mechanisms, from the molecular to the systemic level, exemplifies the elegance with which the brain ensures our energy needs are met in a changing environment, highlighting the intricate interplay of biological process that sustains life (Picard et al., 2018). Understanding these processes opens up new vistas in our approach to diet, disease management, and overall health optimization.

The use of a genetically tractable model organism such as Drosophila melanogaster could provide fundamental insights into the potential relationship between TRP channels and metabolic syndrome. This study delves into the function of TRPγ channels, specifically their role in managing lipid and protein levels. Through genetic manipulation, we explored the consequences of TRPγ mutations on the metabolic processes, revealing critical insights into the regulation of lipid storage and breakdown. We found that TRPγ expression in Dh44 neuroendocrine cells in the brain is critical for maintaining normal carbohydrate levels in tissues (Dhakal et al., 2022). Building on this, we hypothesized that TRPγ in Dh44 cells also regulates lipid and protein homeostasis. Our investigation sheds light on the balance between lipogenesis and lipolysis, processes that respectively synthesize and degrade lipids based on the availability of food, and how these processes are affected by TRPγ activity. By examining the metabolic anomalies in trpγ deficient mutants, including alterations in lipid storage and the consequent impact on starvation resistance, this study advances our understanding of metabolic regulation. Furthermore, we uncover the potential therapeutic benefits of lipase and metformin in counteracting the metabolic deficiencies caused by TRPγ mutations. Through detailed gene expression analysis, we identify the downregulation of crucial genes involved in lipid metabolism, offering new perspectives on the genetic underpinnings of nutrient storage regulation. The implications of this research are far- reaching, not only enhancing our comprehension of basic biological functions but also providing avenues for therapeutic interventions in metabolic disorders.

Results

TRPγ mutants exhibit reduced sugar level alongside elevated lipid and protein levels

In order to understand how D. melanogaster regulates the maintenance of major nutrient levels, including carbohydrates, lipids, and proteins, TRP channel mutants were examined as potential candidates. A recent study revealed that among TRP channel mutants, only trpγ mutant displayed only reduced carbohydrate levels (Dhakal et al., 2022), suggesting the involvement of additional TRP channels in regulating major nutrients. D. melanogaster possesses 13 members of the TRP channel family. Flies with mutations in trpM, trpML, and nompC showed high mortality rates with minimal survival, whereas those with mutations in the remaining 10 TRP superfamily genes were viable and healthy when homozygous. Investigation into the lipid and protein levels of other available TRP superfamily mutants under sated condition revealed that, apart from the trpγ mutant, all tested mutants exhibited normal levels (Figure 1A and B). Specifically, the trpγ1 mutant showed 1.3 times higher triacylglycerol (TAG) levels and 1.5 times higher protein levels compared to the control. These metabolic changes were unique and specific to the trpγ mutants among the evaluated TRP superfamily mutants. Previous observations indicated that trpγ mutants had lower cellular sugar levels and stored glycogen (Dhakal et al., 2022). Consistently, trpγ mutants exhibited reduced levels of cellular sugars (glucose and trehalose), stored glycogen, and sugar levels in the hemolymph (Figure 1C–E). In conclusion, it was found that trpγ1 mutants displayed lower carbohydrate levels but higher lipid and protein levels compared to the control.

TRPg mutants exhibit altered carbohydrate, lipid, and protein levels.

(A) TAG level measurement in the whole-body extract from control (w1118) and TRP channel mutant lines (n=9). (B) Protein level measurement in the whole-body extract from control (w1118) and TRP channel mutant lines (n=6). (C) Measurement of total glucose and trehalose levels (µg/mg) in the whole-body extracts of control (w1118) and trpg1 adult males (n=6). (D) Measurement of tissue glycogen levels (µg/mg) in adult control (w1118) and trpg1 males (n=6). (E) Measurement of Hemolymph glucose + trehalose level in the male flies of control (w1118) and trpγ1 (n=6). (F) Measurement of TAG level in adult males to test for rescue of the TAG defect in trpg flies with the UAS-trpg and the trpgG4 or with the trpg genomic transgene (n=6-10). (G) Rescue experiments showing the measurement of protein levels in adult flies with the indicated genotypes (n=6). Comparisons between multiple experimental groups were conducted via single-factor ANOVA coupled with Scheffe’s post hoc test. The asterisks indicate significant differences from the controls (**P<0.01).

Importantly, the increased TAG and protein levels observed in trpγ1 mutants were confirmed with a second trpγ1 allele (trpγG4) (Figure 1F and G). Furthermore, we restored the deficiency in lipid and protein levels using a genomic trpγ+ transgene, g(trpγ), and by introducing a UAS-trpγ cDNA, controlled by GAL4 knocked into the trpγ locus (trpγG4 flies; +).

TRPγ regulates lipid metabolism through Dh44 neuroendocrine cells in the pars intercerebralis

We recently proposed that TRPγ expression in the six DH44 neuroendocrine cells in the pars intercerebralis (PI) located in the dorsal medial area of the brain (Figure 2–figure supplement 1A1–A3) is essential to maintain normal carbohydrate levels in tissues (Dhakal et al., 2022). Therefore, we next sought to determine whether protein and lipid levels can be regulated by the same neuroendocrine cells. To test this, we used a transgenic fly encoding an inwardly rectifying K+ channel (UAS-Kir2.1) to inactivate specific neurons (Hodge, 2009). Interestingly, inactivating the Dh44 neurons significantly increased the lipid levels but not protein levels (Figure 2A and B). Approximately 16 neurons in the PI express dILP2, which are also overlapped by two DH44-positive neurons in 5 to 10 days old adult male flies (Figure 2–figure supplement 1B1–B3) (Ohhara et al., 2018). However, inactivation of dILP2 neurons did not impair either lipid or protein levels (Figure 2A and B). In conclusion, our findings indicated that DH44 neuroendocrine cells contribute to lipid regulation but not protein regulation.

Dh44 neurons are essential for regulating lipid tissue levels.

(A) Tissue TAG level measurement in whole-body extracts of adult male flies after silencing of Dh44- GAL4 and dILP2-GAL4 with UAS-Kir2.1 (n=6). (B) Measurement of tissue protein level in whole-body extracts of adult male flies. Brain-specific Dh44-GAL4 and dILP2-GAL4 neurons were ablated using UAS-Kir2.1 (n=6). (C) Measurement of TAG from whole- body extract of adult male flies in the indicated genotypes (n=6-7). (D) Measurement of tissue protein level from the whole-body extracts of adult males (n=6). (E) TAG levels in whole-body extracts after RNAi knockdown of trpγ mutants under control of the Dh44- GAL4 (n=6). (F) Measurement of area (µm2) of LDs in adult fat body across the indicated genotypes involved the selection of the 30 largest LDs, choosing the top 10 LDs from each sample for analysis (n=3). (G–L) Nile red stating of fat body from the male of indicated genotypes. The scale bar represents 50 µm. All values are reported as means ± SEM. Comparisons between multiple experimental groups were conducted via single-factor ANOVA coupled with Scheffe’s post hoc test. The asterisks indicate significant differences from the controls (*P<0.05, **P<0.01). Each dot indicates the distribution of individual sample values. (+) and (-) indicate the presence or absence of the indicated transgenes, respectively.

To assess whether trpγ expression in Dh44 neurons is sufficient to restore normal lipid and protein levels in trpγ mutant flies, we expressed the UAS-trpγ under the control of Dh44-GAL4 in the trpγ1 mutant background (Figure 2C and D). Our findings indicated that the increased lipid levels in the trpγ1 mutant background decreased to normal levels via the expression of the trpγ transgene in the DH44 neuroendocrine cells but not in its parent strains (Figure 2C). Again, the expression of the trpγ transgene in the dILP2 neurons had no appreciable effects (Figure 2C). In contrast, the expression of the trpγ transgene did not decrease the protein levels in either the DH44 or the dILP2 neurons (Figure 2D). This indicates that TRPγ expressed in Dh44 cells is sufficient for the regulation of lipid level. Next, RNAi-mediated knockdown experiments were conducted to further examine the role of trpγ in Dh44 neurons. Interestingly, trpγ knockdown in Dh44 neurons significantly increased lipid levels, whereas flies harboring only the Dh44-GAL4 or the UAS-trpγRNAi transgenes displayed normal lipid levels (Figure 2E). This indicated that TRPγ in Dh44 cells is needed for the regulation of lipid levels. Furthermore, the total TAG level in the trpγ1 flies was higher in both males and females (Figure 2–figure supplement 1C), meaning that the functions of trpγ are not sex-specific.

In D. melanogaster, lipids are mainly stored in the form of TAG and cholesterol ester in the adipose tissue [i.e., fat bodies (FBs)] as lipid droplets (LDs) (Liu and Huang, 2013). Consistent with the increased TAG in tissues, Nile red staining of the FBs of trpγ1 and trpγG4 flies exhibited larger lipid mass compared to control animals under sated condition (Figure 2F–I). These LDs returned to their normal size through the expression of trpγ in trpγ or Dh44 neurons but not dILP2 neurons (Figure 2F and J-L). Additionally, inactivating the Dh44 cells (Dh44-GAL4/UAS-Kir2.1) recapitulated the enlarged LD phenotype of the trpγ mutant flies (Figure 2–figure supplement 1D–G). Overall, our findings indicated that trpγ expression regulates lipid and carbohydrate homeostasis but not protein levels in DH44 neuroendocrine cells in the PI.

TRPγ mutants exhibit starvation susceptibility and deficits in lipolysis

Higher lipid cellular levels may decrease lifespan (Johnson and Stolzing, 2019; Tatar et al., 2014). Therefore, we measured the lifespan of control and trpγ1 flies fed with a standard cornmeal diet (Figure 3A). However, the lifespans of control and trpγ1 flies were not significantly different under normal conditions. The LT50 of the control was 59.34 ± 0.92 days and that of trpγ1 was 55.75 ± 2.35 days. In contrast, trpγ appeared to be required for proper metabolism under starvation conditions, as demonstrated by the decreased starvation resistance of the trpγ1 and trpγRNAi knockdown flies in Dh44 cells (Figure 3B) (Dhakal et al., 2022). This defect was fully recovered by the expression of UAS-trpγ under the control of Dh44-GAL4 (Figure 3B).

trpγ1 have deficits in the lipolytic pathway under starvation conditions.

(A) Survival assay to measure the total survival time (days) of control (w1118) and trpγ1 male flies fed with a normal corn meal diet (n=7-8). (B) Survival assay to measure the survival time (h) of the indicated genotypes with male flies under starvation conditions (n=4). (C) TAG level measurement in control (w1118) and trpγ1 adult male flies in both sated (0 h starvation) and starved (24 h starvation) conditions (n=8). (D–E) Nile red staining of the LDs extracted from FB of (D) w1118 and (E) trpγ1 flies under sated (D1 and E1) and starved (D2 and E2) conditions, respectively. The scale bar represents 50 µm. (F) Measurement of area (µm2) of the LDs extracted from the FBs of w1118 and trpγ1 flies under sated (0 h starvation) and starved (24 h starvation) conditions (n=3). (G) Measurement of TAG level with controls and the flies after expressing UAS-AMPKTD under the control of Dh44-GAL4 in the trpγ1 mutant background (n=4). All values are reported as means ± SEM. Survival curves in A and B were estimated for each group, using a Kaplan-Meier method and compared statistically using the log-rank tests. Comparisons between multiple experimental groups in C, F, G, and H were conducted via single-factor ANOVA coupled with Scheffe’s post hoc test. The asterisks indicate significant differences from the controls (**P<0.01).

The starvation sensitive phenotype in the trpγ mutants may have been due to decreased carbohydrate storage in tissues, including glucose, trehalose, and glycogen. However, we previously demonstrated that trpγ mutants could utilize carbohydrates under starved conditions (Dhakal et al., 2022). Furthermore, the elevated whole-body TAG levels in the trpγ mutant flies might suggest that they were unable to break down stored lipids even under starving conditions. Metazoans must coordinate the metabolism of glycogen, lipid, and protein to maintain metabolic homeostasis during fasting periods, thus providing an appropriate energy supply across tissues. Therefore, the total TAG levels of the control and trpγ mutant flies were assessed under sated and starved (starvation for 24 h) conditions (Figure 3C–F. When the control flies were deprived of food for 24 h, their TAG levels decreased significantly (Figure 3C). In contrast, the trpγ1 mutants exhibited no changes in whole-body TAG levels before and after starvation (Figure 3C). To further confirm that lipolysis was restricted in the trpγ mutants under starvation conditions, LDs were stained, after which we measured the sizes of individual LDs accumulated throughout the whole fat bodies of the flies (Figure 3D–F). Despite considerable variations, the sizes of the LDs in the control flies were significantly reduced after starvation. In contrast, no significant differences in LD sizes were identified between the sated and starved conditions in the trpγ1 flies. We examined how overexpression and knockdown of trpγ in Dh44 neurons affect the starvation phenotype. Overexpressing trpγ in Dh44 cells resulted in similarity to the wild-type in both sated and starved conditions, as well as normal survival time under starvation conditions (Figure 3B and Figure 3–figure supplement 1A). Conversely, trpγRNAi knockdown flies in Dh44 neurons reproduced phenotypic traits observed in trpγ mutants, including decreased lipid levels (Figure 3–figure supplement 1B) and reduced survival time under starvation conditions (Figure 3B).

Dh44 neurons regulate starvation-induced sleep suppression (Oh and Suh, 2023), which implies that these neurons become more active under starved conditions. Adenosine monophosphate-activated protein kinase (AMPK) serves as the master controller for maintaining energy balance in cells, coordinating metabolic pathways (Hardie et al., 2003). It regulates the balance between building up and breaking down substances to ensure cellular stability during metabolic stress. AMPK is a key target for treating metabolic diseases like type 2 diabetes and obesity, as its activation increases fatty acid oxidation (Kim et al., 2016). We wonder if activating AMPK in Dh44 neurons improve lipolysis. The expression of AMPKTD, activated form of AMPK (Lee et al., 2007) in Dh44 neurons indeed restored the elevated TAG levels observed in trpγ1 (Figure 3G). This suggests that AMPK functions as a downstream component of Dh44 neuronal activation.

Recovery of the starvation resistance of the TRPγ mutant via metformin treatment

Metformin is widely used to treat many metabolic diseases such as type II diabetes (Lv and Guo, 2020). We previously demonstrated that 1 to 5 mM metformin can induce hypoglycemia, in addition to suppressing fat storage in flies (Nath and Lee, 2024; Sang et al., 2021). Therefore, we next sought to test whether the increased lipid levels of the trpγ mutant can be reduced to a normal level via oral administration of metformin. First, we measured TAG levels at 0, 7, and 14 days after dietary administration of 1 mM or 5 mM of metformin (the treatments were prepared by mixing the appropriate metformin concentrations into standard cornmeal diets) (Figure 4A and B). The high TAG levels in the trpγ mutant (41.91 ±4.56) were decreased to levels similar to those of the control at 7 and 14 days (30.03 ±4.34 and 33.79 ±3.20, respectively) after treatment with 1 mM metformin (Figure 4A). Furthermore, although oral administration of 1 mM metformin did not affect the TAG level in the control, the 5 mM metformin treatment significantly reduced the TAG level of the control flies at 14 days (Figure 4B). Moreover, the 5 mM metformin treatment was more effective in reducing the TAG level in the trpγ mutant. These findings indicated that oral administration of metformin can either suppress lipogenesis or enhance lipolysis in D. melanogaster. To further confirm the effects of metformin treatment, LDs were analyzed under the same condition (Figure 4C and D). The sizes of the LDs in the control and trpγ mutant flies were significantly reduced after treatment with 1 mM metformin at 14 days or 5 mM metformin at 7 and 14 days (Figure 4D).

Rescue of starvation susceptibility phenotype using the lipolytic drug metformin.

(A) TAG level measurement at 0, 7, and 14 days in the control (w1118) and trpγ1 adult male flies after dietary exposure to 1 mM metformin (n=6). (B) TAG measurement at 0, 7, and 14 days in w1118 and trpγ1 adult male flies after dietary exposure to 5 mM metformin (n=6). (C) Pictures of Nile red staining of LDs after dietary exposure to 1 mM and 5 mM metformin in standard fly food for 7 and 14 days in w1118 and trpγ1 flies. The scale bars represent 50 µm. (D) Measurement of area (µm2) of LDs extracted from the FB of w1118 and trpγ1 flies after dietary exposure to 1mM and 5 mM metformin in standard fly food for 7 and 14 days (n=3). (E) Survival assay to measure the survival time (h) of control (w1118) and trpγ1 males after dietary exposure to 1 mM metformin in 1% agar food (n=6-10). (F) Starvation survival assay of control (w1118) and trpγ1 males after dietary exposure to 5 mM metformin in 1% agar food (n=6). All values are reported as means ± SEM. Comparisons between multiple experimental groups were conducted via single-factor ANOVA coupled with Scheffe’s post hoc test. Each dot indicates the distribution of individual sample values. Survival curves in E and F were estimated for each group, using a Kaplan-Meier method and compared statistically using the log-rank tests. The asterisks indicate significant differences from the controls (**P<0.01).

Next, we sought to assess whether metformin could increase starvation resistance in the control and trpγ mutant flies. To test this hypothesis, the starvation resistance of the flies was measured with or without metformin. Neither of the metformin concentrations appeared to affect the starvation resistance of the control flies (Figure 4E and F). However, although metformin did not restore the starvation resistance of the trpγ mutant flies to the level of the controls, it did significantly extend their survival (Figure 4E and F; LT50 of the control at 1 mM and 5 mM: 56.06 ±1.40 h and 52.39 ±2.09 h, respectively; LT50 of the trpγ1 mutant at 1 mM and 5 mM: 50.35 ±3.34 h and 46.42 ±1.54 h, respectively). Our findings thus demonstrated that oral metformin administration can rescue TRPγ-mediated metabolic syndrome.

Downregulation of the lipolytic gene brummer in trpγ mutants

The trpγ1 mutants exhibited marked alterations in the levels of major nutrients. Therefore, we next sought to analyze the transcriptional levels of genes related to gluconeogenesis, lipogenesis, and lipolysis. Gluconeogenesis is a metabolic process that produces glucose from non-carbohydrate carbon substrates. The genome of D. melanogaster harbors two gluconeogenic genes: fructose-1,6-bisphosphatase (fbp) and phosphoenolpyruvate carboxykinase 1 (pepck1) (Miyamoto and Amrein, 2019). The results of our real-time quantitative reverse transcription PCR (qRT-PCR) analyses indicated that the transcription levels of fbp and pepck1 were not significantly different between control flies under sated and starved conditions (Figure 5A). Next, we analyzed two lipogenic genes: acetyl–CoA carboxylase (acc) and desaturase 1 (desat1) (Figure 5A) (Wang et al., 2022; Yee et al., 2013). No significant differences were identified between the control and trpγ1 flies under both conditions. Finally, we investigated the expression of the lipolytic gene brummer (bmm) (Figure 5A) (Blumrich et al., 2021). Under starved conditions, the transcriptional level of bmm was highly increased in the control. In contrast, bmm was downregulated in trpγ1 under sated and starved conditions. Therefore, the transcriptional regulation of w1118 and trpγ1 during starvation was markedly and significantly different. The data indicates that the trpγ mutant can sense the starvation state but responds abnormally by suppressing lipolysis instead of activating it. This dysregulated lipolytic response likely increases the mutant’s vulnerability to starvation, as it cannot effectively mobilize lipid stores for energy during periods of nutrient deprivation. Brummer, a homolog of human adipocyte triglyceride lipase (ATGL), is associated with lipid storage in the form of LDs (Grönke et al., 2005; Men et al., 2016). These results suggest that the lipolytic process is not sufficient to deplete stored lipids in the trpγ1 mutant, resulting in excessive lipid storage. Therefore, the survival time of the trpγ1 mutant during starvation was significantly shorter than that of the control.

Quantitative analysis of the expression of gluconeogenic, lipogenic, and lipolytic genes and the effect of bmm expression and metformin feeding.

(A) qRT-PCR analysis to measure the expression of gluconeogenic genes (fbp, pepck1), lipogenic genes (acc, desat1), and a lipolytic gene (bmm) under sated (0 h starvation) and starved (24 h starvation) conditions in w1118 and trp(1 flies. The relative fold change in the expression of starvation-induced genes (gluconeogenic genes: fbp and pepck1; lipogenic genes: acc and desat1; and lipolytic gene: bmm) was determined in whole- body samples of male adult w1118 and trpγ1 flies by qRT-PCR. Tubulin was used as an internal control to standardize the samples. Each graph shows the number of evaluated samples (n=3). (B) Fluorescence microscopic imaging of bmm::GFP expression in the FBs of w1118 and trpγ1 under sated (0 h starvation) and starvation (24 h starvation) conditions. The scale bar represent 50 µm. (C) Quantification of intensity level of bmm::GFP in the FBs of w1118 and trpγ1 under sated and starvation conditions (n=3). (D–I) Nile red staining of lipids in the FB of flies with the indicated genotypes. The scale bar represent 50 µm. (J) Measurement of area of LDs from samples D–I entailed selecting a total of 30 LDs, with the 10 largest LDs chosen from each sample for analysis (n=3). (K) qRT-PCR analysis to measure the expression of acc, desat1, and bmm from the whole body samples of flies after feeding 5 mM metformin for 1 day (n=3). All values are reported as means ± SEM. Comparisons between multiple experimental groups were conducted via single-factor ANOVA coupled with Scheffe’s post hoc test. The asterisks indicate significant differences from the controls (*P<0.05, **P<0.01).

Transgenic flies expressing bmm::GFP (bmm promoter GFP) in their FB cells were used to examine bmm expression in LDs (Men et al., 2016). The fluorescent images obtained from the transgenic flies were then used to confirm our qRT-PCR results. Two images of control flies under sated and starved conditions were first analyzed both qualitatively and quantitatively (Figure 5B and C). Next, trpγ1 was analyzed under the two conditions (Figure 5B and C). A reduction in fluorescent intensity was clearly detectable in the trpγ1 mutant using the transgenic reporter line. These results further confirmed that trpγ1 has aberrant regulation in the lipolytic pathway.

To further substantiate the idea that the reduction in the expression level of bmm is the primary cause of the enlarged LDs in trpγ mutant flies, we measured LDs in the FB after expressing UAS-bmm using specific GAL4 drivers: trpγG4, Dh44-GLA4, r4- GAL4 (fat body specific), Myo1A-GAL4 (gut enterocyte specific), and Dmef-GAL4 (muscle specific). Only trpγG4, Dh44-GAL4 and r4-GAL4 were able to restore the enlarged LDs defects in the FB, as well as TAG levels of whole body, after expressing UAS-bmm (Figure 5D–J and Figure 5–figure supplement 1). These findings suggest that the downregulation of bmm is one of the factors contributing to the elevated lipid defects in trpγ1. Furthermore, the expression of bmm in the fat body, as well as Dh44 neurons in the PI region, can promote lipolysis at the systemic level.

To conclusively explore whether metformin targets enzymes involved in lipid metabolism, specifically lipolytic or lipogenic genes, we administered 5 mM metformin supplemented in normal food for one day and measured the expression levels of bmm, acc, and desat1. Oral supplementation of metformin resulted in an increase in the expression level of the lipolytic gene, bmm lipase of the control as well as trpγ1 (Figure 5K). Importantly, there was no observed effect of metformin on the expression levels of lipogenic genes. This data provides insights into the mode of action of metformin, suggesting a specific impact on lipolysis.

Recovery of the TRPγ mutant through oral lipase administration or lipid absorption

The intestine is a key organ for lipid absorption and metabolism (Han et al., 2020) and disturbances in intestinal lipid metabolism are associated with hyperlipidemia (Jia et al., 2021). To clarify if our observations were due to insufficient activity of lipase in the gut, lipase enzyme mixed with 1% agar food was supplied to both control and mutants flies. Given that the brief activity of lipase in the intestine, lipase-containing agar food was provided to the flies every 12 h. Surprisingly, supplementation with lipase, but not denatured lipase, under starvation conditions increased the survival of trpγ mutant flies to levels comparable to those of the control flies (Figure 6A; LT50 of lipase treated- control and trpγ1: 50.35 ±0.93 h and 44.94 ±1.33 h, respectively). To support the hypothesis that the extended lifespan of mutant flies during starvation is associated with dietary lipase intake, we examined lipid accumulation (Figure 6–figure supplement 1A–E). Lipase treatment resulted in the breakdown of lipid deposits in the mutant R2 region of intestine. Importantly, only active lipase showed an enhanced survival time under starvation conditions in mutant flies, while denatured lipase did not exhibit this effect. Therefore, our findings suggest that lipase may play a role in digesting accumulated lipid in the intestine, though the mechanisms of lipase absorption remain unclear.

Dietary supply of lipase, glycerol, mixed TAG, and free FA to rescue the starvation sensitivity phenotype.

(A) Survival assays of w1118 and trpγ1 flies under starvation condition by feeding 0.1% lipase or 0.1% denatured lipase mixed into 1% agar food (n=8). (B) Starvation survival assay to measure the survival time (h) of w1118 and trpγ1 male flies after feeding 1% glycerol mixed into 1% agar food (n=4). (C) Survival assay to measure the survival time (h) of w1118 and trpγ1 flies after feeding 0.2% hexanoic acid (HA) supplemented into 1% agar food (n=8). (D) Survival assay to measure the survival time (h) of w1118 and trpγ1 flies after feeding 0.2% mixed (mono-, di-, and tri-) glycerides (n=4). (E–I) Nile red staining of LDs in full gut and the magnified R2 region of adult male flies. (E) w1118, (F) trpγ1, (G) trpγ1,UAS-trpγ/trpγG4, (H) trpγ1,UAS- trpγ/trpγ1;Dh44-GAL4/+. The scale bar represents 50 µm. The arrow indicates the orientation of intestine from anterior to posterior. (I) Measurement of area of LDs in the R2 region of trpγ1, area of 28 LDs from 10 samples (n=10). Note that control and rescued flies have no LD. All values are reported as means ± SEM. Survival curves were estimated for each group, using a Kaplan-Meier method and compared statistically using the log-rank tests. (**P<0.01).

Next, we sought to assess the involvement of each of the components of TAG in the recovery of starvation susceptibility. TAG is composed of glycerol and three fatty acids. Therefore, we first fed the flies with 1% glycerol only. Glycerol extended the survival of the control and trpγ1 flies by 3–4-fold (Figure 6B). However, the LT50 of the control and trpγ1 were still significantly different (Figure 6C; LT50 of control and trpγ1: 244.5 ± 20.10 h and 117.75 ± 3.32 h, respectively). Next, the flies were fed with 0.2% and 0.5% hexanoic acid (HA) (Figure 6C and Figure 6–figure supplement 1F; 0.2% HA, LT50 of control and trpγ1: 65.44 ± 1.94 h and 52.32 ±2.05 h, respectively; 0.5% HA, LT50 of control and trpγ1: 56.27 ±1.30 h and 46.27 ±1.47 h, respectively). Additionally, we also tested the effects of 0.2% and 0.5% concentrations of a TAG mixture (combination of mono-, di-, and tri- acylglycerol) (Figure 6D and Figure 6–figure supplement 1G; 0.2% TAG, LT50 of control and trpγ1: 52.93 ±5.12 h and 50.1 ±3.74 h, respectively; 0.5% TAG, LT50 of control and trpγ1: 65.1 ±2.6 h and 57.75 ±3.32 h, respectively). The HA and mixed TAG treatments significantly extended the survival of the trpγ1 mutants, albeit not to levels similar to those of the control. This indicated that the TRPγ mutants could not fully absorb and burn the digested lipid. Furthermore, lipid-only dietary supplementation markedly limited the LT50 of the control. Importantly, supplementation with lipolytic drugs, lipase, TAG, and free fatty acids effectively rescued the survival of trpγ mutants under starvation conditions. This observation was consistent with our finding that trpγ mutants are unable to degrade lipid stores under starvation conditions. The dietary rescue of trpγ mutants suggests that although the trpγ mutation increases the levels of stored lipids, the flies were not able to utilize them when starved. We next sought to determine whether trpγ mutation reduces lipase production in the abdomen by comparing lipid accumulation in the intestine. The Drosophila midgut is divided into R1–R5 regions, each with distinct functions in nutrient processing and absorption. Notably, the R2 region serves as a critical interface between the fly’s diet and metabolic processes (Capo et al., 2019). To assess lipid distribution, we conducted Nile red staining in control and mutant animals (Figure 6E–H). Mutant flies displayed elevated lipid deposits on the intestinal wall of the R2 region (Figure 6F), a phenotype restored to normal levels by expressing trpγ in its own cells or Dh44 neurons (Figure 6G–I). Consequently, we quantified lipid droplets specifically within the intestinal wall of the R2 region in trpγ1 flies, with no such accumulation observed in control or rescued flies (Figure 6I). These findings indicate that TRPγ plays a crucial role in the brain-gut axis, specifically in controlling lipid metabolism in the intestine, as evidenced by the lipid accumulation observed in the gut region of trpγ1. Taken together, these observations indicate that TRPγ plays an important role in maintaining systemic lipid levels through proper expression of triglyceride lipase.

Assessment of the role of DH44 and its receptors (DH44R1 and DH44R2) in relation to the utilization of fat

In Drosophila, two receptors have been identified for DH44 (Hector et al., 2009). In order to investigate the potential roles of DH44 in fat utilization and identify the receptor responsible for nutrient regulation, we examined lipid and protein levels in Dh44Mi, Dh44R1Mi, and Dh44R2Mi. Dh44Mi and Dh44R1Mishowed normal total TAG levels, while Dh44R2Mi exhibited higher lipid levels like trpγ1 (Figure 7A). Additionally, Nile red staining of LDs in the FBs revealed that Dh44Miand Dh44R2Mi mutants possessed larger LDs (Figure 7B–F), although protein levels remained unaffected across all three mutants (Figure 7G). Further investigations quantified the expression of the bmm in the FBs of Dh44Mi, Dh44R1Mi, and Dh44R2Mi under both sated and starved conditions. In sated conditions, bmm expression was found to be downregulated in Dh44R2Micompared to the control (Figure 7H). Interestingly, under starved conditions, while bmm levels significantly increased in the control, indicating a normal response to starvation, such an increase was not observed in Dh44Mi and Dh44R2Mi. Instead, in these mutants, bmm levels significantly decreased compared to their levels under sated conditions (Figure 7H). In addition, bmm levels in Dh44R1Mi under starved condition did not increase as significantly as in the control. This suggests a unique role of DH44 and its receptors in regulating lipid metabolism and response to nutritional status in Drosophila.

Functional analysis of Dh44, Dh44R1, and Dh44R2 mutants and their roles in lipid accumulation.

(A) Total TAG level (µg TAG/mg fly) measurement in whole body extracts from control (w1118), Dh44Mi, Dh44R1Mi, and Dh44R2Mi mutants (n=5). (B–E) Nile red stains of the fat body from w1118, Dh44Mi, Dh44R1Mi, and Dh44R2Mi respectively. Scale bars represent 50 µm. (F) Area of LDs in each indicated genotype (n=3). (G) The protein (µg protein/mg fly) measurement in the whole-body extracts from control (w1118) and Dh44Mi, Dh44R1Mi, and Dh44R2Mi mutants (n=4). (H) Quantification (qRT-PCR) of lipolytic gene (bmm) expression level in the fat body of w1118, Dh44Mi, Dh44R1Mi, and D44R2Mi flies under sated and starved conditions (n=3). (I) Total TAG level measurement in whole body extracts from the indicated genotypes (n=4). (J–M) Nile red stains of the fat body from (J) w1118, (K) trpγ1 (L) trpγ1,UAS-trpγ1/trpγ1,Dh44R1-GAL4, and (M) trpγ1,UAS-trpγ1/trpγ1,Dh44R2-GAL4. (N) Area of LDs in each indicated genotype (n=3). Scale bar represent 50 µm. (O–Q) Immunohistochemistry with anti- GFP and anti-DsRed. (O and P) Co-expression of trpγ and Dh44R2 (trpγG4/Dh44R2- LexA;UAS-mCD8::GFP/LexAopmCherry) in the R2 region of the inetestine. Arrows indicate co-expression of trpγ and Dh44R2 in the gut cell. (O) Full gut image. Scale bar represents 300 µm. (P) Magnified view of boxed R2 region in O. Scale bar represent 50 µm. (Q) Co-expression of Dh44R2 and Dh44 (Dh44R2-GAL4/Dh44-LexA;UAS- mCD8::GFP/LexAop-mCherry) in the brain. Arrows indicate coexpressed two cells in the PI. Scale bars represent 50 µm. Means ±SEMs. Single factor ANOVA with Scheffe’s analysis was used as a post hoc test to compare multiple sets of data. The asterisks indicate significance from control (*P < 0.05, **P < 0.01). Each dot indicates distribution of individual sample value.

To further verify whether Dh44, Dh44R1, and Dh44R2 mutants can be recovered by oral lipase administration or lipid absorption, we measured the starvation survival time with 1% agar and 1% agar with 0.1% lipase, 1 % glycerol, 0.2 % HA, and 0.2 % mix TAG (Figure 7–figure supplement 1A–D). Under starvation condition, Dh44Miand Dh44R2Mi showed significantly reduced survival time compared to control, whereas Dh44R1Mi flies were normal (LT50 of control, Dh44Mi, Dh44R1Mi, and Dh44R2Mi: 54.32 ± 2.29 h, 36.95 ± 1.76 h, 46.83 ± 1.27 h, and 36.33 ±1.03 h respectively). Interestingly, the reduced survival time observed in Dh44Mi and Dh44R2Mi mutants were improved with the supplementation of lipase (LT50 of Dh44Miand Dh44R2Mi: 54.72 ± 0.55 h and 48.63 ± 1.18 h), glycerol (LT50 of Dh44Miand Dh44R2Mi: 239 ± 1.91 h and 238.5 ± 5.12 h), hexanoic acid (LT50 of Dh44Miand Dh44R2Mi: 62.50 ± 1.50 h and 58.77 ± 0.81 h), and mixed TAG (LT50 of Dh44Mi and Dh44R2Mi: 54.70 ± 3.53 h and 59.50 ± 2.06 h) (Figure 7–figure supplement 1A–D).

To explore the connection between TRPγ and DH44 signaling, we expressed UAS-trpγ under the control of Dh44R1-GAL4 and Dh44R2-GAL4. Only Dh44R2-GAL4 restored the lipid defect in trpγ1, unlike Dh44R1-GAL4 (Figure 7I). Furthermore, Dh44R2-GAL4, but not Dh44R1-GAL4, rescued the excessive lipid accumulation in the FB of trpg1 mutants (Figure 7J–N). Given that Dh44R2 is predominantly expressed in the intestine, we performed immunostaining to examine whether Dh44R2 co-localizes with trpg in gut cells. Our results confirmed that Dh44R2 and trpg are co-expressed in intestinal cells (Figure 7O and P). Additionally, we analyzed Dh44R2 expression in the brain and found that two Dh44R2-expressing cells are co-localized with Dh44- expressing cells in the PI region (Figure 7Q). To further delineate whether Dh44R2- mediated fat utilization is specific to the brain, gut, or fat body, we knocked down Dh44R2RNAiusing Dh44-GAL4, Myo1A-GAL4, and cg-GAL4, respectively (Figure 7–figure supplement 1E). Notably, knockdown of Dh44R2 with Myo1A-GAL4 resulted in elevated TAG levels, indicating that DH44R2 activity in lipid metabolism is specific to the gut.

Discussion

Here, we found that the function of TRPγ, one of the TRPC channels of D. melanogaster, in DH44 neuroendocrine cells plays an essential role in lipid regulation, which is linked to alterations in membrane lipids (Cobb et al., 2021; Overgaard et al., 2005). The trpγ1 mutant exhibited clear signs of metabolic syndrome, as demonstrated by reduced carbohydrate level coupled with much higher protein and lipid levels in the body (Figure 8). More importantly, we found that the expression of trpγ in DH44 neurons was necessary and sufficient to regulate the carbohydrate and lipid. However, we failed to identify the specific cells required for regulating protein homeostasis because the increased protein levels in the trpγ mutants were not recovered by pilot screening (data not shown) except in the trpγ-GAL4. In conclusion, our findings suggested that the DH44 system has important roles in regulating the metabolic homeostasis of carbohydrates and lipids (Figure 8).

Schamatic model representing the roles of trpγ in metabolic regualtion of carbohydrate, lipid, and protein.

The metabolic dysregulation of carbohydrates and lipids observed in the trpγ mutants was phenocopied by the cellular inactivation of DH44 cells. Compared to the controls, the Dh44-GAL4/UAS-Kir2.1 flies exhibited lower carbohydrate levels and increased lipid levels. We previously proposed that TRPγ holds DH44 neurons in a state of afterdepolarization, thus reducing firing rates by inactivating voltage-gated Na+ channels (Dhakal et al., 2022). At the physiological level, this induces the consistent release of DH44 and depletion of DH44 stores, resulting in nutrient utilization and storage malfunctions. Likewise, our findings revealed that Dh44 mutant and its receptor mutant, Dh44R2, displayed defects analogous to those observed in the trpγ mutant. The trpγ mutant phenotype in lipid regulation can be restored by the expression of trpγ in the Dh44-GAL4 as well as Dh44R2-GAL4. This indicates that TRPγ function in the brain and gut independently for lipid homeostasis. The mode of action for TRPγ in the gut should be further investigated. Finally, it would also be interesting to investigate the potential roles of TRPC members in humans, especially the involvement of TRPC4 and TRPC5 in metabolic syndrome.

Multiple lines of evidence have suggested that DH44 neurons communicate with the intestine. First, DH44 at the PI in the brain is an incretin-like hormone that acts as a hemolymph nutrient sensor. The secretion of DH44 is highly dependent on the calorie value of a given diet, as DH44 secretion is only mediated by nutritional sugars but not non-nutritional sugars (Dhakal et al., 2022; Dus et al., 2015; Oh et al., 2021). Second, the expression of TRPγ in the PI is sufficient to recover any metabolic defects such as reduced carbohydrate and increased lipid levels, as well as related phenomena such as starvation resistance. DH44 and kinin neuropeptides influence desiccation and starvation tolerance in D. melanogaster (Cannell et al., 2016). There is an inconsistency of total TAG levels and the LD size observed in the Dh44 mutant. This inconsistency raises a possibility that the signaling pathway from DH44 release to its receptor DH44R2 only accounts for part of the lipid metabolic process under starvation. While Dh44 mutant flies displayed normal internal TAG levels, Dh44R2 mutant flies exhibited elevated TAG levels. This suggested that the lipolysis phenotype could be facilitated by a neuropeptide other than DH44. Alternatively, a DH44 neuropeptide-independent pathway could mediate the lipolysis. Although Dh44Mi exhibited normal TAG levels, the presence of larger LDs indicated that DH44 plays a regulatory role in lipid metabolism.

LDs consist of a hydrophobic core of neutral storage fats, such as triglycerides or cholesterol esters, encased by a protein-coated phospholipid monolayer. From our findings, it is suggested that DH44 and its receptor DH44R2, but not DH44R1, are involved in the control of lipid storage, indicating a specific pathway through which DH44 influences lipid metabolism distinct from its role in carbohydrate metabolism and overall energy homeostasis. TRPg in DH44 neurons may influence the release of metabolic signals or hormones that act on postsynaptic DH44R2 cells. These postsynaptic cells could, in turn, modulate lipid storage and metabolism in a non-cell autonomous manner. However, the mechanism by which TRPg functions in DH44R2 cells remains unclear. One possible explanation is that TRPg in the gut may be activated by stretch or osmolarity (Akitake et al., 2015). Third, trpg gene knockdown through RNAi in the DH44 neurons is enough to phenocopy all of the metabolic defects observed in the trpg mutants. Fourth, trpg mutants exhibit low expression of brummer lipase, which results in accumulation of LDs in the body, which is a reciprocal process regulated by DH44 neurons. Brummer lipase is essential for regulating lipid levels in the insect fat body by mediating lipid mobilization and energy homeostasis. In Nilaparvata lugens, it facilitates triglyceride breakdown (Lu et al., 2018), while studies in Drosophila show that reduced Brummer lipase expression decreases fatty acids and increases diacylglycerol levels, highlighting its role in lipid metabolism (Nazario-Yepiz et al., 2021). Here, we additionally demonstrate that bmm expression in DH44 neurons within the PI region can systemically regulate TAG levels. Cell signaling or energy status in DH44 neurons may contribute to hormonal release that targets organs such as the fat body.

Fifth, the activation of AMPK in the DH44 neurons can recover the defects of lipid level of trpγ1. Sixth, the reduced starvation resistance can be significantly recovered through dietary lipase supplementation. These findings support the assertion that the sugar- sensing DH44 neurons send downward signals to the intestine by absorbing the carbohydrates and lipids derived from dietary intake. However, DH44 neurons do not regulate protein levels in the body because the increased protein levels of trpγ mutants were not recovered by the expression of trpγ in the DH44 neurons. These findings confirm that trpγ regulates bmm expression in the adult bowel, enabling dietary lipid digestion and maintaining systemic lipid homeostasis.

Treatment with metformin, a lipolytic agent, at two different concentrations (1 mM and 5 mM) decreased the total TAG levels and reduced the size of the LDs in D. melanogaster. Similarly, metformin increased the starvation resistance of the trpγ1 mutants but not the controls. These findings were consistent with the widely acknowledged anti-obesity properties of metformin (Sang et al., 2021). Furthermore, our results were in agreement with the evidence that metformin helps to reduce lipid accumulation caused by diets rich in saturated fatty acids (Kim et al., 2020). Although the exact target of metformin is not known, its lipolytic effect is significant and widely documented. Oral administration of 5 mM of metformin reduced normal TAG levels of wild-type flies, whereas the 1 mM dose had no significant effect. Therefore, low metformin doses were effective at decreasing stored lipid levels in flies. The accumulation of intracellular LDs is regulated by autophagy. Metformin improves autophagic flux through AMPKα-mediated mechanisms (Kim et al., 2020). In the Drosophila context, we propose a mechanism involving AMPK activation within DH44 neurons as a rescue mechanism for starvation resistance in trpγ1, highlighting its crucial role in this process. AMPK activation is known to enhance an organism’s ability to endure environmental stresses by facilitating autophagy-mediated cellular cleanup, promoting mitochondrial biogenesis for increased energy production, and optimizing metabolic pathways for enhanced energy efficiency (Kim et al., 2020). While the precise nature of the connection between TRPγ, DH44 neuronal activity, and AMPK activation remains unspecified, it is conceivable that TRPγ influences AMPK activation or regulation within DH44 neurons, contributing to stress responses, ATP supply, and metabolic adaptations. Metformin can influence lipid metabolism through various mechanisms, one of which is the promotion of lipolysis (Szkudelski et al., 2022). This effect results in a gradual reduction of stored lipid levels over time. Orally administering metformin to flies led to a significant reduction in lipid levels, accompanied by an increase in survival. Subsequent qPCR analysis post-metformin feeding revealed a significant upregulation in the expression level of the lipolytic gene (bmm). These results strongly support our hypothesis that metformin induces lipolysis when orally administered.

The exact mechanism through which orally administered lipase leads to gastric lipolysis remains unclear. However, our experimental data indicated that 0.1% lipase increased the survival time of trpγ1 under starvation. Lipase exhibits a high degree of chemical selectivity and has the ability to catalyze triglyceride into glycerol ester, monoglycerides, glycerol, and fatty acids (Liu et al., 2020). Free fatty acids, HA, and glycerol at concentrations of 0.2%–0.5% can be utilized as energy sources and increase survival time under starvation conditions, suggesting that trpγ1 participates in the utilization of HA and glycerol under starvation conditions.

The therapeutic properties of metformin as a lipolytic and anti-diabetic drug have been widely documented in humans. However, additional studies are needed to identify the exact target of metformin. In this context, our findings demonstrate that D. melanogaster can be an excellent model to study the mode of action and biosafety of metformin.

Materials and methods

Key resources table

Chemical sources

Trehalase from porcine kidney (cat. # 9025-52-9), amyloglucosidase (cat. # 9032-08-0), the glucose (HK) assay kit (cat. # GAHK-20-1KT), and triglyceride mix (cat. # 17810- 1AMP-S), lipase from Candida rugosa (cat. # OT705690000), metformin (cat. # D150959), Nile Red (cat. # N3013), glycerol (cat. # G5516) were purchased from Sigma-Aldrich Co. Hexanoic acid (cat. # W255912) were purchased from Wako Pure

Chemical Industry Ltd. The PierceTM BCA protein assay kit (cat. # 23225) and the LiquiColor triglyceride test kit (cat. # 2100-225) were purchased from Thermo Fisher Scientific and Stanbio Laboratory, respectively.

TAG level measurement

TAG levels were quantified as described previously with some modifications using a LiquiColor Triglyceride Test kit (cat. # 2100-225, Stanbio Laboratory, Germany) (Sang et al., 2021). Ten 5–10 day-old male flies were weighed and crushed in 1 mL of PBST (1X PBS and 0.2 percent Triton X-100). The homogenate was incubated at 70 °C for 5 minutes and centrifuged for 3 minutes at 9,500 g. Next, 100 µL of supernatant was transferred into a 1.5 mL Eppendorf tube and mixed with 1 mL of Stanbio LiquiColor Triglyceride Test kit reagent or 1 mL of deionized water to provide a baseline. The reaction mixture was kept at 37 °C for 15 minutes. Finally, the absorbance of the sample solution was measured at 500 nm using a spectrophotometer and the TAG level was calculated based on a standard calibration curve.

Protein level measurement

Protein assays were performed as previously described using the PierceTM BCA Protein Assay Kit with some modifications. Briefly, ten 5–10 day-old male flies were weighed and crushed in 1 mL of PBST (1X PBS and 0.2 percent Triton X-10) and incubated at 70 °C for 5 minutes. The homogenate was then centrifuged for 3 minutes at 9,500 g, after which 300 µL of supernatant was mixed with 600 µL of PierceTM BCA Protein Assay Kit (UF289330). After a 30-minute incubation period at 37 °C, the absorbance of the samples was measured at 530 nm using a spectrophotometer and compared to a standard calibration curve for quantification.

Trehalose and glucose measurements in whole adult flies or hemolymph

Whole-body glucose and trehalose levels were measured in adult flies as previously described (Meunier et al., 2007). Briefly, ten 5–10 day old males were collected, weighed, and crushed in 250 µL of 0.25 M Na2CO3 buffer. The homogenates were then incubated in a water bath (95°C) for 5 min to inactivate all enzymes. Next, 600 µL of 0.25 M sodium acetate and 150 µL of 1 M acetic acid (pH 5.2) were added to the samples, after which the mixtures were centrifuged at 12,500 g for 10 min at 24 °C. Afterward, 200 µL of supernatant was transferred a new microfuge tube and 2 µL porcine kidney trehalase (Sigma: T8778 UN) was added to the sample mixture and incubated overnight at 37 °C to convert trehalose into glucose. Next, 1 mL of glucose hexokinase solution (Sigma: GAHK-20) was added to 100 µL of the sample and incubated for 20 min at 37 °C. Optical density (OD) values were measured at 340 nm. Finally, total glucose and trehalose levels were calculated using a standard glucose curve generated through similar reactions with standard trehalose and glucose.

Glucose and trehalose level measurement in hemolymph were extracted from the male flies. Flies were punctured in the thorax with the help of fine injection needle and kept in 0.5 mL tubes having a punctured base with 21-gauge needle. The punctured flies were adjusted by keeping the shoulder down to prevent leakage from the genital tract. The tubes were set into the 1.5 mL microfuge tube and centrifuged at 4°C for 5 minutes at 2800 g force. 0.5 µL of hemolymph was added to 14.5 µL PBS. 2 µL of porcine kidney trehalase (Sigma: T8778 UN) was added to the sample mixture and incubated overnight at 37°C to convert trehalose into glucose. 1 mL of glucose hexokinase solution (Sigma: GAHK-20) was added to the 100 µL of the sample and incubated at 37°C for 20 minutes. Optical density (OD) values were measured at 340 nm. Total glucose and trehalose levels were calculated using a standard glucose curve generated through similar reactions with standard trehalose and glucose.

Glycogen measurements

Tissue glycogen levels from whole-body extracts of adult flies were quantified as previously described (Dus et al., 2011). Briefly, 5 males were homogenized in 100 µL of ice-chilled 1X PBS after measuring the weight. The homogenates were kept at 70 °C for 5 min to inactivate all metabolic enzymes. The homogenates were then centrifuged at 12,500 g for 3 min at 4 °C, after which 20 µL of the supernatant was transferred to 1.5 mL microfuge tubes and diluted with 1X PBS to a 1:3 ratio. An amyloglucosidase dilution was prepared by mixing 1.5 µL of amyloglucosidase (Sigma A1602) suspension in 998.5 µL 1X PBS. Finally, a 20 µL aliquot of the diluted sample was added to 20 µL of the diluted amyloglucosidase solution, as well as to the glycogen standard. Both the glycogen standard and test samples were then incubated at 37 °C for 1 h. A commercial glucose (HK) assay reagent (Sigma: G3293 VER) was used to measure total glucose at 340 nm. The glycogen level was quantified by comparing it with standard curve plotted from standard glycogen samples.

Survival assay

Survival experiments were conducted as previously described (Lee et al., 2018). Normal cornmeal diets were used to conduct the normal survival assays. To measure survival under starvation conditions, twenty 3–4-day-old male flies were fed with 1% agar food supplemented with or without various concentrations of lipase (0.1%; active or denatured), glycerol (1%), hexanoic acid (0.2%, 0.5%), glycerol (1%), TAG mix (0.2%, 0.5%), or metformin (1 mM, 5 mM). Every 12 hours, the flies were monitored/counted and then transferred to fresh vials with the same food supply. The experiments were conducted until the food vials had been cleared.

Immunohistochemistry

We performed immunohistochemistry as previously described (Dhakal et al., 2022). Briefly, to fix and block the specimens, we placed freshly dissected tissues (brain and intestine) into a well of 24-well tissue culture plate (Costar Corp.) placed on ice, which contained 940 μL of fixing buffer (1 mM EGTA, 0.1M Pipes pH 6.9, 2 mM MgSO4, 1% TritonX-100, 150 mM NaCl). 60 μL formaldehyde (37%) was added to the wells and mixed instantly before the tissues were added. As many as tissues that were dissected within 15 minutes were fixed and then the samples were incubated for 30 min more. Samples were washed with wash buffer (1X PBS, 0.1% saponin) three times (15 min. for each washing), and blocked with 1 mL blocking buffer (1X PBS, 0.1% saponin, and 5 mg/mL BSA) at 4 °C for 4-8 hours.

To perform immunostaining, primary antibodies were inserted into the sample at 4°C for 18 hours (Jais and Brüning, 2022). Samples were washed with wash buffer three times (15 min each) and added secondary antibodies {(1:200) goat anti-mouse Alexa Fluor 488 (cat. # A11029), goat anti-mouse Alexa Fluor 568 (cat. # A11004), goat anti- rabbit Alexa Fluor 488 (cat. # A11034), and goat anti-rabbit Alexa 568 (cat. # A11036) at 4 °C for 4 hours. Finally, samples were washed three times and stored into 1.25X PDA (187.5 mM NaCl, 37.5% glycerol, 62.5 mM Tris pH 8.8), and kept at 4°C for more than 1 hr. Samples were mounted and examined using a Leica Stellaris 5 Confocal Microscope.

Nile red staining

Nile Red is a dark purplish-red powder (Sigma N-3013), the stock solution must be prepared in acetone (1000 µg/mL) and kept in a tightly sealed, lightproof container at 4 °C. Briefly, 5–10 day-old male flies were fixed in a sagittal position on a glass slide and submerged in a 1 X PBS solution. FBs were gently dissected from a dorsal abdominal region along with thorax or the gut under a stereomicroscope. The dissected tissues were fixed with 4% formaldehyde solution for 15 minutes at room temperature. The fixed tissues were gently washed three times with 1X PBS (5 min for each wash). Nile Red (1:1000 dilution) were added to the tissue samples for 5 minutes. Finally, the stained tissues were washed with 1 mL 1X PBS and mounted in 50% glycerol on a glass slide. LD deposition exhibits greater density in the upper abdominal region compared to the lower abdominal area. This comparison was meticulously conducted within the identical segments 2-3 of the abdomen to ensure accuracy.

For intestinal lipid staining, 5-10 days old male flies were fixed on a glass slide and submerged in 1X PBS solution. Full intestine was dissected very carefully. The dissected intestine was fixed with 4% formaldehyde solution for 15 minutes at room temperature. The fixed intestine was washed 3 times with 1X PBS. Nile red was added to the sample for 5 minutes. The stained samples were washed with 1mL 1X PBS and mounted in 50% glycerol on a glass slide. The stained tissues were examined using a Leica Stellaris 5 Confocal Microscope.

qRT-PCR

qRT-PCR assays were performed as described previously (Yoshinari et al., 2021). Briefly, ten 6–10 day-old male flies (control and mutant) were selected for whole body and 15 flies were selected for fat body samples in Supplement Figure 6K. The experiments were conducted under sated (0 h starvation) and starved (24 h starvation) conditions. Total RNA was extracted using the TRizol reagent (Invitrogen) followed by DNase (Promega) treatment. cDNA was synthesized using the AMV reverse transcriptase system (Promega). RT-qPCR experiments were carried out using a Bio- rad CFX system. The Takara TB Green Premix was used to assess the mRNA expression level of each gene according to the manufacturer’s instructions. Relative gene expression was calculated using the 2−ΔΔCt method. Three biological samples were used and transcript levels were normalized to the D. melanogaster housekeeping gene tubulin. The experiments were conducted using the following primer pairs: acc, 5’-ACG AGG GCG AGC AGC GTT AC-3′ (forward) and 5′-TAG GGC GAC TTG GTG GGC AT-3′ (reverse); bmm, 5′-ATG ACT TCG GAC TTC TTC AGG G-3′ (forward) and 5′-CCA ATT CAG ATG GAA GAG CTG-3′ (reverse); fbp, 5′-CTC CAA CGA GCT GTT CAT CA-3′ (forward) and 5′-TGA ACC GAT CGA CAC CAG GC-3′ (reverse); pepck1, 5′-AGG TGC ACA TCT GCG ATG GC-3′ (forward) and 5′-CCA CCA CGT AAG CAG AGT CC-3′ (reverse); desat1, 5’-AAG CCG GTG CCC AGT CCA TC-3’ (forward) and 5’-ATG GTC GCG AGC CCA ATG GT-3’ (reverse); and tubulin, 5′-TCC TTG TCG CGT GTG AAA CA-3′ (forward) and 5′-CCG AAC GAG TGG AAG ATG AG-3’ (reverse).

Statistics and Reproducibility

D. melanogaster was selected as a model organism in this study. For the experiments, male flies were mostly used unless we did not mention the sex. All of the experiments were conducted under laboratory conditions. The appropriate number of replicates was established based on previous research. A large enough sample size was used in all of our assays to ensure that our results were representative and repeatable. No data points were left out of the analysis. For each genotype, the data points indicate the values of individual replicates. The error bars in all of the figures represent the standard error of the mean (SEM). The analysis of the RT-qPCR data was conducted using the CT values. Comparisons between multiple experimental groups were conducted via single-factor ANOVA and Scheffe’s post hoc test. Pair-wise comparisons were conducted via Student’s t-test. Survival curves were estimated for each group, using a Kaplan-Meier method and compared statistically using the log-rank tests. The asterisks in the figures indicate statistical significance (*P<0.05, **P<0.01). All statistical analyses were conducted using the Origin Pro 8 software for Windows (ver. 8.0932; Origin Lab Corporation, USA).